\documentclass[9pt]{article}

% Package for setting margins
\usepackage[margin=0.5in]{geometry}

% Additional useful packages
\usepackage{amsmath, amssymb, amsthm}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Lecture Notes: MATH 447},
    pdfpagemode=FullScreen,
}
\usepackage{breqn}

% Theorem Styles
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{note}{Note}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}


\title{MATH 447: Real Variables \\ Lecture Notes}
\author{Compiled by Jerich Lee}
\date{\today}

\begin{document}
\maketitle

\section*{Peano Axioms}
\begin{definition}[Peano Axioms]
The natural numbers \( \mathbb{N} \) are defined by the following postulates:
\begin{enumerate}[label=(N\arabic*)]
    \item \( \mathbb{N} \) contains a distinguished element \( 1 \).
    \item Every \( n \in \mathbb{N} \) has its successor in \( \mathbb{N} \), denoted \( S(n) \).
    \item \( 1 \) is not the successor of any element in \( \mathbb{N} \).
    \item If \( m \) and \( n \) have the same successor, then \( m = n \).
    \item If \( A \subseteq \mathbb{N} \) such that \( 1 \in A \) and \( S(n) \in A \) whenever \( n \in A \), then \( A = \mathbb{N} \).
\end{enumerate}
\end{definition}

\begin{theorem}[Uniqueness of \( \mathbb{N} \)]
Suppose \( X \) is a set with a distinguished element \( 1' \) and a successor map \( S' \), satisfying the Peano Axioms (N1-N5). Then there exists a bijection \( \Phi : \mathbb{N} \to X \) such that:
\begin{align*}
    \Phi(1) = 1', \quad \Phi(S(n)) = S'(\Phi(n)) \, \forall n \in \mathbb{N}.
\end{align*}
\end{theorem}

\section*{Mathematical Induction}
\begin{theorem}[Principle of Mathematical Induction]
Suppose \( (P_n)_{n \in \mathbb{N}} \) is a sequence of statements such that:
\begin{enumerate}
    \item \( P_1 \) is true.
    \item For any \( n \in \mathbb{N} \), \( P_n \) implies \( P_{n+1} \).
\end{enumerate}
Then \( P_n \) is true for all \( n \in \mathbb{N} \).
\end{theorem}

\begin{example}[Induction Proof]
Prove \( \sum_{k=1}^n k = \frac{n(n+1)}{2} \) for \( n \in \mathbb{N} \).

\textbf{Proof.} Base case (\( n=1 \)):
\[
\sum_{k=1}^1 k = 1 = \frac{1(1+1)}{2}.
\]
Inductive step: Assume \( \sum_{k=1}^n k = \frac{n(n+1)}{2} \). Then:
\[
\sum_{k=1}^{n+1} k = \sum_{k=1}^n k + (n+1) = \frac{n(n+1)}{2} + (n+1).
\]
Simplify:
\[
\sum_{k=1}^{n+1} k = \frac{n(n+1) + 2(n+1)}{2} = \frac{(n+1)(n+2)}{2}.
\]
Thus, \( \sum_{k=1}^n k = \frac{n(n+1)}{2} \) holds for all \( n \).
\end{example}

\section*{Properties of Integers}
\begin{definition}[Addition Properties of \( \mathbb{Z} \)]
The integers \( \mathbb{Z} \) satisfy:
\begin{enumerate}[label=(A\arabic*)]
    \item \textbf{Associativity:} \( a + (b + c) = (a + b) + c \) for all \( a, b, c \in \mathbb{Z} \).
    \item \textbf{Commutativity:} \( a + b = b + a \) for all \( a, b \in \mathbb{Z} \).
    \item \textbf{Neutral Element:} \( \exists 0 \in \mathbb{Z} \) such that \( a + 0 = a \).
    \item \textbf{Existence of Opposites:} For every \( a \in \mathbb{Z} \), \( \exists -a \in \mathbb{Z} \) such that \( a + (-a) = 0 \).
\end{enumerate}
\end{definition}

\begin{theorem}[Uniqueness of Additive Elements]
\begin{enumerate}
    \item The neutral element \( 0 \) is unique.
    \item For any \( a \in \mathbb{Z} \), the opposite \( -a \) is unique.
\end{enumerate}
\end{theorem}

\begin{proof}
1. Suppose \( 0 \) and \( 0' \) are both neutral elements. Then:
\[
0 = 0 + 0' = 0'.
\]
2. Suppose \( a + x = 0 \) and \( a + y = 0 \). Then:
\[
x = x + 0 = x + (a + y) = (x + a) + y = 0 + y = y.
\]
Thus, \( -a \) is unique.
\end{proof}
\section*{Properties of \( \mathbb{Z} \) and \( \mathbb{Q} \)}
\subsection*{Properties of Addition on \( \mathbb{Z} \)}
\begin{definition}[Addition Properties of \( \mathbb{Z} \)]
\leavevmode
\begin{enumerate}[label=(A\arabic*)]
    \item \textbf{Associativity:} \( a + (b + c) = (a + b) + c \) for all \( a, b, c \in \mathbb{Z} \).
    \item \textbf{Commutativity:} \( a + b = b + a \) for all \( a, b \in \mathbb{Z} \).
    \item \textbf{Neutral Element:} \( \exists 0 \in \mathbb{Z} \) such that \( a + 0 = a \) for all \( a \in \mathbb{Z} \).
    \item \textbf{Existence of Opposites:} \( \forall a \in \mathbb{Z}, \exists -a \in \mathbb{Z} \) such that \( a + (-a) = 0 \).
\end{enumerate}
\end{definition}

\begin{proposition}[Uniqueness of \( 0 \) and \( -a \)]
\begin{enumerate}
    \item The neutral element \( 0 \) is unique.
    \item For each \( a \in \mathbb{Z} \), the opposite \( -a \) is unique.
\end{enumerate}
\end{proposition}

\subsection*{Multiplication on \( \mathbb{Z} \)}
\begin{definition}[Multiplication Properties of \( \mathbb{Z} \)]
\leavevmode
\begin{enumerate}[label=(M\arabic*)]
    \item \textbf{Associativity:} \( a \cdot (b \cdot c) = (a \cdot b) \cdot c \) for all \( a, b, c \in \mathbb{Z} \).
    \item \textbf{Commutativity:} \( a \cdot b = b \cdot a \) for all \( a, b \in \mathbb{Z} \).
    \item \textbf{Neutral Element:} \( \exists 1 \in \mathbb{Z} \) such that \( 1 \cdot a = a \) for all \( a \in \mathbb{Z} \).
    \item \textbf{Distributive Law:} \( (a + b) \cdot c = a \cdot c + b \cdot c \) for all \( a, b, c \in \mathbb{Z} \).
\end{enumerate}
\end{definition}

\begin{proposition}[Multiplication by Zero]
For any \( a \in \mathbb{Z} \), \( 0 \cdot a = 0 \).
\end{proposition}

\subsection*{Properties of \( \mathbb{Q} \)}
\begin{definition}[Field Properties of \( \mathbb{Q} \)]
The rational numbers \( \mathbb{Q} \) satisfy:
\begin{enumerate}[label=(M\arabic*), resume]
    \item \textbf{Inverse:} \( \forall a \in \mathbb{Q} \setminus \{0\}, \exists a^{-1} \in \mathbb{Q} \) such that \( a \cdot a^{-1} = 1 \).
\end{enumerate}
If \( (X, +, 0, \cdot, 1) \) satisfies (A1–A4), (M1–M4), and the distributive law, \( X \) is called a field.
\end{definition}

\subsection*{Ordered Fields}
\begin{definition}[Ordered Fields]
A field \( F \) is ordered if equipped with a linear order \( \leq \) such that:
\begin{enumerate}[label=(O\arabic*)]
    \item If \( a \leq b \), then \( a + c \leq b + c \) for all \( a, b, c \in F \).
    \item If \( a \leq b \) and \( c \geq 0 \), then \( ac \leq bc \).
\end{enumerate}
\end{definition}

\begin{theorem}[Properties of Ordered Fields]
Let \( F \) be an ordered field. Then for all \( a, b, c \in F \):
\begin{enumerate}[label=(\roman*)]
    \item If \( a \leq b \), then \( -b \leq -a \).
    \item If \( a \leq b \) and \( c \leq 0 \), then \( bc \leq ac \).
    \item If \( 0 \leq a \) and \( 0 \leq b \), then \( 0 \leq ab \).
    \item \( 0 \leq a^2 \) for all \( a \in F \).
\end{enumerate}
\end{theorem}

\subsection*{Rational Zeros Theorem}
\begin{theorem}[Rational Zeros Theorem]
Suppose \( p(x) = c_nx^n + \ldots + c_1x + c_0 \), with \( c_0, \ldots, c_n \in \mathbb{Z} \), \( c_0 \neq 0 \), \( c_n \neq 0 \). If \( p(r) = 0 \) for \( r = \frac{c}{d} \) (where \( c, d \in \mathbb{Z} \), \( d \neq 0 \), \( \gcd(c, d) = 1 \)), then \( c \mid c_0 \) and \( d \mid c_n \).
\end{theorem}

\begin{corollary}[Irrationality of \( \sqrt{2} \)]
No rational number \( r \) satisfies \( r^2 = 2 \).
\end{corollary}
\section*{Ordered Fields and Completeness}

\subsection*{Fields and Order}
\begin{proposition}
If \( F \) is a field with more than one element, then \( 0 \neq 1 \).
\end{proposition}
\begin{proof}
Let \( x \in F \) be distinct from \( 0 \). Then \( 0 = x \cdot 0 \neq x \cdot 1 = x \), hence \( 0 \neq 1 \).
\end{proof}

\begin{definition}[Ordered Fields]
A field \( F \) is called ordered if it is equipped with a linear order \( \leq \) satisfying:
\begin{enumerate}[label=(O\arabic*)]
    \item If \( a \leq b \), then \( a + c \leq b + c \) for all \( a, b, c \in F \).
    \item If \( a \leq b \) and \( c \geq 0 \), then \( ac \leq bc \).
\end{enumerate}
\end{definition}

\subsection*{Properties of Ordered Fields}
\begin{theorem}
Let \( F \) be an ordered field. Then for all \( a, b, c \in F \):
\begin{enumerate}[label=(\roman*)]
    \item If \( a \leq b \), then \( -b \leq -a \).
    \item If \( a \leq b \) and \( c \leq 0 \), then \( bc \leq ac \).
    \item If \( 0 \leq a \) and \( 0 \leq b \), then \( 0 \leq ab \).
    \item \( 0 \leq a^2 \) for all \( a \in F \).
\end{enumerate}
\end{theorem}

\subsection*{Absolute Value and Distance}
\begin{definition}[Absolute Value]
For \( a \in F \), the absolute value \( |a| \) is defined as:
\[
|a| =
\begin{cases}
a & \text{if } a \geq 0, \\
-a & \text{if } a < 0.
\end{cases}
\]
\end{definition}

\begin{definition}[Distance]
The distance between \( a, b \in F \) is defined as:
\[
\text{dist}(a, b) = |a - b|.
\]
\end{definition}

\subsection*{Completeness Axiom}
\begin{definition}[Completeness Axiom]
If \( S \subset \mathbb{R} \) is non-empty and bounded above, then it has a unique least upper bound (supremum), denoted \( \sup S \).
\end{definition}

\subsection*{Archimedean Property and Denseness of \( \mathbb{Q} \)}
\begin{proposition}[Archimedean Property]
If \( a, b > 0 \) in \( \mathbb{R} \), then there exists \( n \in \mathbb{N} \) such that \( n \cdot a > b \).
\end{proposition}

\begin{theorem}[Denseness of \( \mathbb{Q} \)]
The rational numbers \( \mathbb{Q} \) are dense in \( \mathbb{R} \), meaning that for any \( a, b \in \mathbb{R} \) with \( a < b \), there exists \( r \in \mathbb{Q} \) such that \( a < r < b \).
\end{theorem}
\section*{Sequences and Limits (Sections 7-9)}

\subsection*{Definitions and Examples}
\begin{definition}[Sequence]
A sequence is a function \( s : \{m, m + 1, \dots \} \to \mathbb{R} \) (for \( m \in \mathbb{Z} \)). We denote the sequence as \( (s_n)_{n \geq m} \), where \( s_n = s(n) \).
\end{definition}

\begin{definition}[Convergence and Limit]
A sequence \( (s_n) \) converges to \( L \in \mathbb{R} \) if:
\[
\forall \varepsilon > 0, \exists N \in \mathbb{R} \text{ such that } |s_n - L| < \varepsilon \text{ for } n > N.
\]
We write \( \lim_{n \to \infty} s_n = L \) or \( s_n \to L \).
\end{definition}

\begin{proposition}[Uniqueness of Limits]
A sequence cannot have more than one limit.
\end{proposition}

\subsection*{Examples of Limits}
\begin{enumerate}[label=\arabic*.]
    \item \( \lim_{n \to \infty} \frac{1}{n} = 0 \).
    \item The sequence \( ((-1)^n)_{n \in \mathbb{N}} \) does not converge.
    \item \( \lim_{n \to \infty} \frac{3n + 1}{2n + 1} = \frac{3}{2} \).
\end{enumerate}

\subsection*{Facts about Limits}
\begin{proposition}
If \( (s_n) \) converges, and \( s_n \geq a \) for all but finitely many \( n \), then \( \lim s_n \geq a \).
\end{proposition}

\begin{proposition}
If \( s_n \geq 0 \) for all \( n \) and \( \lim s_n = s \), then \( \lim \sqrt{s_n} = \sqrt{s} \).
\end{proposition}

\subsection*{Convergent Sequences are Bounded}
\begin{definition}[Bounded Sequence]
A sequence \( (s_n) \) is called bounded if \( \exists A \in \mathbb{R} \) such that \( |s_n| \leq A \) for all \( n \).
\end{definition}

\begin{theorem}
Convergent sequences are bounded.
\end{theorem}

\subsection*{Arithmetic of Limits}
\begin{theorem}[Limits of Sums, Products, and Ratios]
Suppose \( \lim s_n = s \) and \( \lim t_n = t \). Then:
\begin{enumerate}
    \item \( \lim (s_n + t_n) = s + t \),
    \item \( \lim (a \cdot s_n) = a \cdot s \) for any \( a \in \mathbb{R} \),
    \item \( \lim (s_n \cdot t_n) = s \cdot t \),
    \item If \( t \neq 0 \), then \( \lim \frac{s_n}{t_n} = \frac{s}{t} \).
\end{enumerate}
\end{theorem}
\section*{Sums, Products, Ratios of Limits (Section 9)}

\subsection*{Arithmetic of Limits}
\begin{theorem}[Arithmetic of Limits]
Suppose \( \lim s_n = s \) and \( \lim t_n = t \). Then:
\begin{enumerate}
    \item \( \lim (s_n + t_n) = s + t \),
    \item \( \lim (a \cdot s_n) = a \cdot s \) for any \( a \in \mathbb{R} \),
    \item \( \lim (s_n \cdot t_n) = s \cdot t \),
    \item If \( t \neq 0 \), then \( \lim \frac{s_n}{t_n} = \frac{s}{t} \).
\end{enumerate}
\end{theorem}

\begin{theorem}[Squeeze Theorem]
If \( a_n \leq s_n \leq b_n \) for all \( n \), and \( \lim a_n = \lim b_n = s \), then \( \lim s_n = s \).
\end{theorem}

\subsection*{Basic Examples of Limits}
\begin{theorem}[Basic Examples]
\leavevmode
\begin{enumerate}
    \item \( \lim \frac{1}{n^p} = 0 \) for \( p > 0 \),
    \item \( \lim a^n = 0 \) if \( |a| < 1 \),
    \item \( \lim n^{1/n} = 1 \),
    \item \( \lim a^{1/n} = 1 \) for \( a > 0 \).
\end{enumerate}
\end{theorem}

\subsection*{Diverging Sequences}
\begin{definition}[Divergence to Infinity]
We say \( \lim s_n = +\infty \) if for all \( A > 0 \), there exists \( N \in \mathbb{R} \) such that \( s_n > A \) for \( n > N \). Similarly, \( \lim s_n = -\infty \) is defined.
\end{definition}

\begin{theorem}[Product Rule for Divergence]
If \( \lim s_n = +\infty \) and \( \lim t_n > 0 \), then \( \lim (s_n \cdot t_n) = +\infty \).
\end{theorem}

\begin{theorem}
If \( s_n > 0 \) for all \( n \), then \( \lim s_n = +\infty \) if and only if \( \lim \frac{1}{s_n} = 0 \).
\end{theorem}

\subsection*{Monotone Sequences (Section 10)}
\begin{definition}[Monotone Sequences]
A sequence \( (s_n) \) is:
\begin{itemize}
    \item \textit{Increasing} if \( s_n \leq s_{n+1} \) for all \( n \),
    \item \textit{Decreasing} if \( s_n \geq s_{n+1} \) for all \( n \),
    \item \textit{Monotone} if it is either increasing or decreasing.
\end{itemize}
\end{definition}

\begin{example}[Examples of Monotone Sequences]
\leavevmode
\begin{enumerate}
    \item \( x_n = \sum_{k=1}^n \frac{1}{k^2} \) is increasing because \( x_{n+1} = x_n + \frac{1}{(n+1)^2} > x_n \).
    \item \( y_n = \frac{(-1)^n}{n^2} \) is not monotone because \( y_{n+1} > y_n \) if \( n \) is odd, and \( y_{n+1} < y_n \) if \( n \) is even.
\end{enumerate}
\end{example}
\section*{Monotone Sequences and Convergence (Section 10)}

\subsection*{Monotone Sequences}
\begin{definition}[Monotone Sequences]
A sequence \( (s_n) \) is:
\begin{itemize}
    \item \textit{Increasing} if \( s_n \leq s_{n+1} \) for all \( n \),
    \item \textit{Decreasing} if \( s_n \geq s_{n+1} \) for all \( n \),
    \item \textit{Monotone} if it is either increasing or decreasing.
\end{itemize}
\end{definition}

\begin{theorem}[Theorem 10.2]
Any monotone bounded sequence converges.
\end{theorem}

\begin{example}[Bounded Monotone Sequence]
Consider \( s_n = \sum_{k=0}^{n-1} \frac{1}{k!} \). This sequence is:
\begin{itemize}
    \item Increasing, since \( s_{n+1} = s_n + \frac{1}{n!} > s_n \).
    \item Bounded, since \( s_n \leq 3 \) (using an induction-based proof that \( k! \geq 2^{k-1} \) for \( k \geq 1 \)).
\end{itemize}
Thus, \( \lim s_n = e \approx 2.71828 \).
\end{example}

\subsection*{Unbounded Monotone Sequences}
\begin{theorem}[Theorem 10.4]
If \( (s_n)_{n \geq m} \) is an unbounded increasing (decreasing) sequence, then \( \lim s_n = +\infty \) (resp. \( \lim s_n = -\infty \)).
\end{theorem}

\begin{example}[Harmonic Sequence]
Let \( s_n = \sum_{k=1}^n \frac{1}{k} \). This sequence is:
\begin{itemize}
    \item Increasing, since \( s_{n+1} = s_n + \frac{1}{n+1} > s_n \).
    \item Unbounded, as shown using a lower bound argument:
    \[
    s_{2^m} \geq \frac{1}{1} + \frac{1}{2} + \cdots + \frac{1}{2^{m-1}} \geq m.
    \]
\end{itemize}
Thus, \( s_n \to +\infty \).
\end{example}

\subsection*{Lim Sup and Lim Inf}
\begin{definition}[Lim Sup and Lim Inf]
Let \( u_N = \sup\{s_n : n > N\} \) and \( v_N = \inf\{s_n : n > N\} \). Then:
\[
\limsup s_n = \lim_{N \to \infty} u_N, \quad \liminf s_n = \lim_{N \to \infty} v_N.
\]
\end{definition}

\begin{theorem}[Properties of Lim Sup and Lim Inf]
\leavevmode
\begin{enumerate}
    \item \( \limsup s_n \geq \liminf s_n \).
    \item If \( \lim s_n \) exists, then \( \limsup s_n = \lim s_n = \liminf s_n \).
    \item If \( \limsup s_n = \liminf s_n = s \), then \( \lim s_n = s \).
\end{enumerate}
\end{theorem}

\begin{example}[Oscillating Sequence]
Let \( s_n = 
\begin{cases} 
\frac{1}{n}, & n \text{ even} \\
-n, & n \text{ odd}
\end{cases}.
\)
Then:
\begin{itemize}
    \item \( \limsup s_n = 0 \),
    \item \( \liminf s_n = -\infty \).
\end{itemize}
\end{example}

\subsection*{Decimal Expansions}
\begin{theorem}
Any real number can be expressed as a decimal expansion \( K.d_1d_2d_3\ldots \), where \( K \in \{0, 1, 2, \dots \} \) and \( d_k \in \{0, \ldots, 9\} \). For instance:
\[
1 = 1.000\ldots = 0.999\ldots
\]
\end{theorem}
\section*{Lim Inf, Lim Sup, and Cauchy Sequences (Sections 10-11)}

\subsection*{Lim Inf and Lim Sup}
\begin{definition}[Lim Sup and Lim Inf]
Let \( u_N = \sup\{s_n : n > N\} \) and \( v_N = \inf\{s_n : n > N\} \). Then:
\[
\limsup s_n = \lim_{N \to \infty} u_N, \quad \liminf s_n = \lim_{N \to \infty} v_N.
\]
\end{definition}

\begin{theorem}[Theorem 10.7]
\leavevmode
\begin{enumerate}
    \item If \( \lim s_n \) is defined, then \( \liminf s_n = \lim s_n = \limsup s_n \).
    \item If \( \liminf s_n = s = \limsup s_n \), then \( \lim s_n = s \).
\end{enumerate}
\end{theorem}

\begin{example}
Let \( s_n = 
\begin{cases} 
\frac{1}{n} & \text{if } n \text{ is even}, \\
-n & \text{if } n \text{ is odd}.
\end{cases}
\)
Then:
\begin{itemize}
    \item \( \limsup s_n = 0 \),
    \item \( \liminf s_n = -\infty \).
\end{itemize}
\end{example}

\subsection*{Cauchy Sequences}
\begin{definition}[Cauchy Sequence]
A sequence \( (s_n) \) is called Cauchy if:
\[
\forall \varepsilon > 0, \exists N \in \mathbb{N} \text{ such that } |s_n - s_m| < \varepsilon \text{ for } n, m > N.
\]
\end{definition}

\begin{theorem}[Theorem 10.11]
A sequence \( (s_n) \) converges if and only if it is Cauchy.
\end{theorem}

\begin{lemma}[Lemma 10.9]
Any convergent sequence is Cauchy.
\end{lemma}

\begin{lemma}[Lemma 10.10]
Any Cauchy sequence is bounded.
\end{lemma}

\subsection*{Subsequences}
\begin{definition}[Subsequence]
A sequence \( (t_k) \) is a subsequence of \( (s_n) \) if there exists a strictly increasing sequence \( n_1 < n_2 < \ldots \) such that \( t_k = s_{n_k} \) for any \( k \).
\end{definition}

\begin{example}[Subsequence Examples]
\leavevmode
\begin{itemize}
    \item \( s_n = \frac{1}{n} \): A subsequence \( t_k = \frac{1}{k^2} \), where \( n_k = k^2 \).
    \item \( s_n = (-1)^n + \frac{1}{n} \): The sequence diverges, but \( s_{2k} = 1 + \frac{1}{2k} \) converges.
\end{itemize}
\end{example}

\begin{theorem}[Subsequential Limits]
Every sequence has a subsequence that converges to a limit.
\end{theorem}
\section*{Subsequences and Subsequential Limits (Section 11)}

\subsection*{Subsequences}
\begin{definition}[Subsequence]
A sequence \( (t_k) \) is a subsequence of \( (s_n) \) if there exists a strictly increasing sequence \( n_1 < n_2 < \ldots \) such that \( t_k = s_{n_k} \) for any \( k \).
\end{definition}

\begin{lemma}[Subsequences of Subsequences]
Any subsequence of a subsequence of \( (s_n) \) is a subsequence of \( (s_n) \).
\end{lemma}

\subsection*{Convergence of Subsequences}
\begin{theorem}[Theorem 11.3]
If \( \lim s_n = s \) (finite or \( \pm\infty \)), then any subsequence \( (t_k) \) has the same limit.
\end{theorem}

\begin{proof}
Let \( \lim s_n = s \), and let \( t_k = s_{n_k} \). Then for \( \varepsilon > 0 \), there exists \( N \) such that \( |s_n - s| < \varepsilon \) for \( n > N \). Since \( n_k \to \infty \), we can find \( K \) such that \( n_k > N \) for \( k > K \). Thus, \( |t_k - s| < \varepsilon \) for \( k > K \), implying \( \lim t_k = s \).
\end{proof}

\subsection*{Monotone Subsequences}
\begin{theorem}[Theorem 11.4]
Every sequence has a monotone subsequence.
\end{theorem}

\begin{corollary}[Bolzano-Weierstrass Theorem]
Every bounded sequence has a convergent subsequence.
\end{corollary}

\begin{example}[Divergent Sequence with Convergent Subsequence]
Let \( s_n = (-1)^n \left(1 + \frac{1}{n}\right) \). This sequence is bounded but divergent. The subsequence \( s_{2k} = 1 + \frac{1}{2k} \) converges to \( 1 \).
\end{example}

\subsection*{Subsequential Limits}
\begin{definition}[Subsequential Limit]
A subsequential limit of \( (s_n) \) is any limit of a subsequence, possibly \( \pm\infty \).
\end{definition}

\begin{theorem}[Theorem 11.2]
Suppose \( (s_n) \) is a sequence.
\begin{enumerate}
    \item \( t \in \mathbb{R} \) is a subsequential limit if and only if \( \forall \varepsilon > 0, \{n : |s_n - t| < \varepsilon\} \) is infinite.
    \item \( t = +\infty \) (or \( t = -\infty \)) is a subsequential limit if \( (s_n) \) is not bounded above (or below).
\end{enumerate}
\end{theorem}

\subsection*{Lim Inf and Lim Sup as Subsequential Limits}
\begin{theorem}[Theorem 11.7]
For any sequence \( (s_n) \), \( \limsup s_n \) and \( \liminf s_n \) are limits of monotone subsequences.
\end{theorem}

\begin{theorem}[Theorem 11.8]
Let \( S \) be the set of subsequential limits of \( (s_n) \). Then:
\begin{enumerate}
    \item \( S \) is non-empty.
    \item \( \inf S = \liminf s_n, \quad \sup S = \limsup s_n \).
    \item \( \lim s_n \) exists if and only if \( S \) consists of a single point, \( S = \{\lim s_n\} \).
\end{enumerate}
\end{theorem}
\section*{Lim Sup, Lim Inf, and Metric Spaces (Sections 11-13)}

\subsection*{Subsequential Limits}
\begin{definition}[Subsequential Limit (Definition 11.6)]
For a sequence \( (s_n) \), a subsequential limit is any limit of a subsequence (in \( \mathbb{R} \cup \{\pm\infty\} \)).
\end{definition}

\begin{theorem}[Properties of Subsequential Limits (Theorem 11.2)]
Suppose \( (s_n) \) is a sequence.
\begin{enumerate}
    \item \( t \in \mathbb{R} \) is a subsequential limit if and only if \( \forall \varepsilon > 0, \{n : |s_n - t| < \varepsilon\} \) is infinite.
    \item \( t = +\infty \) (\( t = -\infty \)) is a subsequential limit if \( (s_n) \) is not bounded above (resp. below).
\end{enumerate}
\end{theorem}

\subsection*{The Set of Subsequential Limits}
\begin{theorem}[Theorem 11.8]
Suppose \( (s_n) \) is a sequence, and \( S \) is the set of subsequential limits. Then:
\begin{enumerate}
    \item \( S \) is non-empty.
    \item \( \inf S = \liminf s_n \) and \( \sup S = \limsup s_n \).
    \item \( \lim s_n \) exists if and only if \( S \) consists of a single point. Then \( \{\lim s_n\} = S \).
\end{enumerate}
\end{theorem}

\begin{corollary}[Convergent Sequences]
If \( \lim s_n = s \), then \( S = \{s\} \), \( \limsup s_n = s = \liminf s_n \).
\end{corollary}

\subsection*{Lim Sup and Lim Inf Revisited}
\begin{theorem}[Theorem 12.1]
If \( \lim s_n = s \in (0, \infty) \), then for any sequence \( (t_n) \):
\[
\limsup (s_n t_n) = s \cdot \limsup t_n.
\]
\end{theorem}

\begin{corollary}[Corollary 12.3]
If \( (s_n) \) is a sequence of positive numbers and \( \lim \frac{s_{n+1}}{s_n} \) exists, then:
\[
\lim s_n^{1/n} \text{ also exists, and } \lim s_n^{1/n} = \lim \frac{s_{n+1}}{s_n}.
\]
\end{corollary}

\begin{example}
\leavevmode
\begin{enumerate}
    \item \( \lim (n!)^{1/n} = +\infty \).
    \item \( \lim \frac{1}{n} (n!)^{1/n} = \frac{1}{e} \).
\end{enumerate}
\end{example}

\subsection*{Metric Spaces}
\begin{definition}[Metric (Definition 13.1)]
A metric \( d : S \times S \to [0, \infty) \) satisfies:
\begin{enumerate}[label=(D\arabic*)]
    \item \textbf{Non-degeneracy:} \( d(x, y) = 0 \iff x = y \).
    \item \textbf{Symmetry:} \( d(x, y) = d(y, x) \) for all \( x, y \in S \).
    \item \textbf{Triangle Inequality:} \( d(x, y) + d(y, z) \geq d(x, z) \) for all \( x, y, z \in S \).
\end{enumerate}
\end{definition}

\begin{example}[Metrics]
\leavevmode
\begin{itemize}
    \item On \( \mathbb{R} \): \( d(x, y) = |x - y| \).
    \item On \( \mathbb{R}^n \): \( d(\vec{x}, \vec{y}) = \sqrt{\sum_{i=1}^n (x_i - y_i)^2} \).
\end{itemize}
\end{example}

\subsection*{Convergence in Metric Spaces}
\begin{definition}[Convergence (Definition 13.2)]
A sequence \( (s_n) \subset S \) converges to \( s \in S \) if:
\[
\lim_{n \to \infty} d(s_n, s) = 0.
\]
\end{definition}

\begin{definition}[Cauchy Sequence (Definition 13.2)]
A sequence \( (s_n) \subset S \) is Cauchy if:
\[
\forall \varepsilon > 0, \exists N \text{ such that } d(s_n, s_m) < \varepsilon \text{ for all } n, m > N.
\]
\end{definition}

\begin{proposition}[Cauchy and Convergence]
If \( (s_n) \) converges in \( (S, d) \), then \( (s_n) \) is Cauchy.
\end{proposition}

\begin{definition}[Complete Metric Spaces]
A metric space \( (S, d) \) is complete if every Cauchy sequence in \( S \) converges to a point in \( S \).
\end{definition}

\begin{example}
The space \( \mathbb{R}^n \) with the Euclidean metric is complete.
\end{example}
\section*{Metric Spaces (Section 13)}

\subsection*{Definition and Examples}
\begin{definition}[Metric (Definition 13.1)]
Suppose \( S \) is a set. A function \( d : S \times S \to [0, \infty) \) is called a metric if the following hold:
\begin{enumerate}[label=(D\arabic*)]
    \item \textbf{Non-degeneracy:} \( d(x, y) = 0 \iff x = y \) (hence \( d(x, y) > 0 \) when \( x \neq y \)).
    \item \textbf{Symmetry:} \( d(x, y) = d(y, x) \) for all \( x, y \in S \).
    \item \textbf{Triangle Inequality:} \( d(x, y) + d(y, z) \geq d(x, z) \) for all \( x, y, z \in S \).
\end{enumerate}
\end{definition}

\begin{example}[Examples of Metrics]
\leavevmode
\begin{itemize}
    \item On \( \mathbb{R} \): \( d(x, y) = |x - y| \).
    \item On \( \mathbb{R}^n \): \( d(\vec{x}, \vec{y}) = \sqrt{\sum_{i=1}^n (x_i - y_i)^2} \) (Euclidean metric).
    \item Discrete Metric: For \( x, y \in S \), define:
    \[
    d(x, y) = 
    \begin{cases} 
    0 & \text{if } x = y, \\
    1 & \text{if } x \neq y.
    \end{cases}
    \]
\end{itemize}
\end{example}

\subsection*{Convergence and Completeness}
\begin{definition}[Convergence (Definition 13.2)]
Suppose \( (S, d) \) is a metric space. A sequence \( (s_n) \subset S \) converges to \( s \in S \) if \( \lim_{n \to \infty} d(s_n, s) = 0 \); that is, \( \forall \varepsilon > 0, \exists N \text{ such that } d(s_n, s) < \varepsilon \text{ for } n > N \).
\end{definition}

\begin{definition}[Cauchy Sequence (Definition 13.2 continued)]
A sequence \( (s_n) \subset S \) is called Cauchy if:
\[
\forall \varepsilon > 0, \exists N \text{ such that } d(s_n, s_m) < \varepsilon \text{ for } n, m > N.
\]
\end{definition}

\begin{proposition}[Cauchy Sequences are Convergent in Complete Spaces]
If \( (S, d) \) is complete, then every Cauchy sequence in \( S \) converges to a point in \( S \).
\end{proposition}

\begin{example}[Completeness of \( \mathbb{R} \)]
\( \mathbb{R} \) with the standard metric \( d(x, y) = |x - y| \) is complete. Any Cauchy sequence in \( \mathbb{R} \) converges to a real number.
\end{example}

\begin{example}[Non-Completeness of \( \mathbb{Q} \)]
Consider \( \mathbb{Q} \) with \( d(x, y) = |x - y| \). The sequence \( r_n \in (\sqrt{2} - \frac{1}{n}, \sqrt{2}) \) is Cauchy in \( \mathbb{Q} \) but does not converge in \( \mathbb{Q} \) because \( \sqrt{2} \notin \mathbb{Q} \).
\end{example}

\subsection*{Special Metrics}
\begin{example}[Manhattan (Taxicab) Metric]
For \( \vec{x} = (x_1, \ldots, x_n) \) and \( \vec{y} = (y_1, \ldots, y_n) \) in \( \mathbb{R}^n \), the taxicab metric is defined as:
\[
d_1(\vec{x}, \vec{y}) = \sum_{i=1}^n |x_i - y_i|.
\]
\end{example}

\begin{proposition}[Completeness of \( (\mathbb{R}^n, d_1) \)]
The metric space \( (\mathbb{R}^n, d_1) \) is complete.
\end{proposition}

\subsection*{Inner Product and Triangle Inequality}
\begin{definition}[Inner Product]
For \( \vec{x}, \vec{y} \in \mathbb{R}^n \), define the inner product:
\[
\langle \vec{x}, \vec{y} \rangle = \sum_{i=1}^n x_i y_i.
\]
The magnitude of \( \vec{x} \) is:
\[
\|\vec{x}\| = \sqrt{\langle \vec{x}, \vec{x} \rangle}.
\]
\end{definition}

\begin{theorem}[Bunyakovsky-Cauchy-Schwarz Inequality]
For all \( \vec{x}, \vec{y} \in \mathbb{R}^n \):
\[
|\langle \vec{x}, \vec{y} \rangle| \leq \|\vec{x}\| \|\vec{y}\|.
\]
\end{theorem}

\begin{lemma}[Triangle Inequality Lite]
For \( \vec{x}, \vec{y} \in \mathbb{R}^n \):
\[
\|\vec{x} + \vec{y}\| \leq \|\vec{x}\| + \|\vec{y}\|.
\]
\end{lemma}
\section*{Metric Spaces: Bounded Sets, Open and Closed Sets, and Closure (Section 13)}

\subsection*{Bounded Sets}
\begin{definition}[Bounded Sets]
A set \( E \) in a metric space \( (S, d) \) is bounded if there exists \( y \in S \) such that:
\[
\sup_{x \in E} d(y, x) < \infty.
\]
\end{definition}

\begin{remark}
If such a \( y \) exists, then for any \( z \in S \), \( \sup_{x \in E} d(z, x) < \infty \). This follows from the triangle inequality:
\[
d(z, x) \leq d(y, x) + d(z, y).
\]
\end{remark}

\begin{example}
A sequence \( (x_k) \) is bounded if the set \( \{x_1, x_2, \ldots\} \) is bounded. That is, for some (or any) \( y \in S \):
\[
\sup_k d(y, x_k) < \infty.
\]
\end{example}

\subsection*{Bolzano-Weierstrass Theorem}
\begin{theorem}[Bolzano-Weierstrass for \( \mathbb{R}^n \)]
Any bounded sequence in \( \mathbb{R}^n \) has a convergent subsequence.
\end{theorem}

\begin{example}[Failure of Bolzano-Weierstrass in Discrete Metrics]
Consider \( \mathbb{N} \) equipped with the discrete metric \( d(x, y) = 1 \) for \( x \neq y \) and \( d(x, y) = 0 \) for \( x = y \). The sequence \( x_n = n \) is bounded but has no convergent subsequences because convergent sequences are eventually constant in discrete metrics.
\end{example}

\subsection*{Interior Points and Open Sets}
\begin{definition}[Open Ball]
An open ball with center \( s_0 \) and radius \( r > 0 \) is:
\[
B_r^o(s_0) = \{s \in S : d(s, s_0) < r\}.
\]
\end{definition}

\begin{definition}[Interior Points]
A point \( s_0 \in S \) is interior to \( E \subset S \) if there exists \( r > 0 \) such that \( B_r^o(s_0) \subset E \). The set of all interior points is denoted \( E^o \), called the interior of \( E \).
\end{definition}

\begin{definition}[Open Sets]
A set \( E \subset S \) is open if \( E = E^o \).
\end{definition}

\begin{example}
\leavevmode
\begin{itemize}
    \item In \( S = \mathbb{R} \) with the usual metric, \( [0, \infty) \) is not open, but \( (0, \infty) \) is.
    \item In \( S = \mathbb{R}^2 \), the set \( E = \{(x, 0) : x \geq 0\} \) has \( E^o = \emptyset \).
\end{itemize}
\end{example}

\subsection*{Properties of Open and Closed Sets}
\begin{theorem}[Facts about Open Sets]
\leavevmode
\begin{enumerate}
    \item \( S \) and \( \emptyset \) are open.
    \item A union of any collection of open sets is open.
    \item A finite intersection of open sets is open.
\end{enumerate}
\end{theorem}

\begin{definition}[Closed Sets]
A set \( E \subset S \) is closed if \( S \setminus E \) is open.
\end{definition}

\begin{theorem}[Facts about Closed Sets]
\leavevmode
\begin{enumerate}
    \item \( S \) and \( \emptyset \) are closed.
    \item An intersection of any collection of closed sets is closed.
    \item A finite union of closed sets is closed.
\end{enumerate}
\end{theorem}

\subsection*{Closure and Boundary}
\begin{definition}[Closure]
The closure of \( E \subset S \), denoted \( \overline{E} \), is the intersection of all closed sets containing \( E \).
\end{definition}

\begin{definition}[Boundary]
The boundary of \( E \subset S \) is:
\[
\partial E = \overline{E} \setminus E^o.
\]
\end{definition}

\begin{example}[Closure in \( \mathbb{R} \)]
Let \( E = \left\{\frac{1}{n} : n \in \mathbb{N}\right\} \subset \mathbb{R} \). Then:
\[
\overline{E} = E \cup \{0\}.
\]
\end{example}
\section*{Closure, Boundary, and Open/Closed Sets (Section 13)}

\subsection*{Open Balls and Open Sets}
\begin{definition}[Open Ball]
For \( s_0 \in S \) and \( r > 0 \), the open ball with center \( s_0 \) and radius \( r \) is:
\[
B_r^o(s_0) = \{s \in S : d(s, s_0) < r\}.
\]
\end{definition}

\begin{proposition}
A set \( E \subset S \) is open if and only if it is a union of open balls.
\end{proposition}

\subsection*{Closed Sets and De Morgan's Laws}
\begin{definition}[Closed Sets]
A set \( E \subset S \) is closed if \( S \setminus E \) is open.
\end{definition}

\begin{proposition}[Properties of Closed Sets]
\leavevmode
\begin{enumerate}
    \item \( S \) and \( \emptyset \) are closed.
    \item Any intersection of closed sets is closed.
    \item A finite union of closed sets is closed.
\end{enumerate}
\end{proposition}

\begin{proposition}[De Morgan's Laws]
For any collection \( \{A_i\}_{i \in I} \subset S \):
\[
S \setminus \bigcup_{i \in I} A_i = \bigcap_{i \in I} (S \setminus A_i), \quad S \setminus \bigcap_{i \in I} A_i = \bigcup_{i \in I} (S \setminus A_i).
\]
\end{proposition}

\subsection*{Examples of Open and Closed Sets}
\begin{example}[Intervals in \( \mathbb{R} \)]
\leavevmode
\begin{itemize}
    \item \( (a, b) \) is open, but not closed.
    \item \( [a, b] \) is closed, but not open.
    \item \( (a, b], [a, b) \) are neither open nor closed.
\end{itemize}
\end{example}

\begin{example}[Discrete Metric]
In a discrete metric space:
\begin{itemize}
    \item Every set is both open and closed.
\end{itemize}
\end{example}

\subsection*{Closure and Boundary}
\begin{definition}[Closure]
The closure of \( E \subset S \), denoted \( \overline{E} \), is the intersection of all closed sets containing \( E \).
\end{definition}

\begin{definition}[Boundary]
The boundary of \( E \subset S \) is:
\[
\partial E = \overline{E} \setminus E^o.
\]
\end{definition}

\subsection*{Properties of Closure and Boundary}
\begin{proposition}
\leavevmode
\begin{enumerate}
    \item \( E = \overline{E} \) if and only if \( E \) is closed.
    \item \( s \in \overline{E} \) if and only if \( s \) is a limit of a sequence in \( E \).
    \item \( \partial E = \overline{E} \cap (S \setminus E)^- \).
\end{enumerate}
\end{proposition}

\begin{example}[Closure of \( E = \left\{\frac{1}{n} : n \in \mathbb{N}\right\} \subset \mathbb{R} \)]
The closure is:
\[
\overline{E} = \left\{\frac{1}{n} : n \in \mathbb{N}\right\} \cup \{0\}.
\]
\end{example}
\section*{Compactness (Section 13)}

\subsection*{Definition of Compactness}
\begin{definition}[Compactness (Definition 13.11)]
Suppose \( E \subset S \). A family \( \mathcal{U} \) of open sets is an \textit{open cover} for \( E \) if:
\[
E \subset \bigcup_{U \in \mathcal{U}} U.
\]
A \textit{subcover} is a subfamily of \( \mathcal{U} \) which is also an open cover. \( E \) is called \textit{compact} if every open cover has a finite subcover.
\end{definition}

\begin{note}
A cover \( \mathcal{U} \) is a collection of sets, not their union. Thus, a cover is a subset of \( \mathcal{P}(S) \) (the power set of \( S \)), not \( S \).
\end{note}

\subsection*{Examples in \( \mathbb{R} \) with Usual Metric}
\begin{enumerate}
    \item \( E = [0, \infty) \) is not compact. For example:
    \begin{itemize}
        \item \( U_k = (-1, k) \) (\( k \in \mathbb{N} \)) is an open cover with no finite subcover.
    \end{itemize}
    \item \( E = (0, 1) \) is not compact. For example:
    \begin{itemize}
        \item \( U_k = (1/k, 1) \) (\( k \in \mathbb{N} \)) is an open cover with no finite subcover.
    \end{itemize}
    \item \( E = [a, b] \) (\( a, b \in \mathbb{R} \)) is compact (proof to follow).
\end{enumerate}

\begin{proposition}[Compactness of Finite Sets]
Any finite set is compact.
\end{proposition}
\begin{proof}
Let \( E = \{e_1, \ldots, e_N\} \). For any open cover \( \mathcal{U} \) of \( E \), select \( U_i \in \mathcal{U} \) containing \( e_i \). Then \( \{U_1, \ldots, U_N\} \) is a finite subcover.
\end{proof}

\subsection*{Compactness and Boundedness}
\begin{proposition}
Any compact set is bounded.
\end{proposition}
\begin{proof}
If \( E \) is not bounded, then for \( s \in S \), the sets \( B_k^o(s) \) (\( k \in \mathbb{N} \)) form an open cover of \( E \) with no finite subcover.
\end{proof}

\begin{example}[Bounded but Not Compact]
Equip \( \mathbb{N} \) with the discrete metric \( d(x, y) = 
\begin{cases} 
0 & x = y, \\
1 & x \neq y.
\end{cases} \)
Then \( \mathbb{N} \) is bounded, but it is not compact because the open cover \( U_n = \{n\} \) (\( n \in \mathbb{N} \)) has no finite subcover.
\end{example}

\subsection*{Properties of Compact Sets}
\begin{proposition}
\leavevmode
\begin{enumerate}
    \item A closed subset of a compact set is compact.
    \item A finite union of compact sets is compact.
\end{enumerate}
\end{proposition}

\subsection*{Nested Sequences of Closed Sets}
\begin{proposition}
Suppose \( F_1 \supset F_2 \supset \cdots \) are closed non-empty subsets of a compact set \( E \). Then:
\[
\bigcap_{n} F_n \neq \emptyset, \quad \text{and it is compact.}
\]
\end{proposition}

\subsection*{Heine-Borel Theorem and Cantor Set}
\begin{theorem}[Heine-Borel Theorem]
A subset of \( \mathbb{R}^n \) is compact if and only if it is closed and bounded.
\end{theorem}

\begin{example}[Cantor Set]
Define:
\[
F_0 = [0, 1], \quad F_1 = [0, 1/3] \cup [2/3, 1], \quad F_2 = [0, 1/9] \cup [2/9, 1/3] \cup [2/3, 7/9] \cup [8/9, 1], \dots
\]
The Cantor set \( C = \bigcap_n F_n \) is non-empty, closed, and compact. It contains no intervals, and its interior is empty.
\end{example}
\section*{Compactness and Total Boundedness (Section 13)}

\subsection*{Definition of Compactness}
\begin{definition}[Compactness (Definition 13.11)]
Suppose \( E \subset S \). A family \( \mathcal{U} \) of open sets is an \textit{open cover} for \( E \) if:
\[
E \subset \bigcup_{U \in \mathcal{U}} U.
\]
A \textit{subcover} is a subfamily of \( \mathcal{U} \) which is also an open cover. \( E \) is called \textit{compact} if every open cover has a finite subcover.
\end{definition}

\begin{note}
A cover \( \mathcal{U} \) is a collection of sets, not their union. Thus, a cover is a subset of \( \mathcal{P}(S) \) (the power set of \( S \)), not \( S \).
\end{note}

\subsection*{Compactness and Completeness}
\begin{proposition}
Suppose \( E \subset S \). If \( E \) is compact, then \( E \) is complete.
\end{proposition}

\begin{proof}
Suppose \( E \) is not complete. Then there exists a Cauchy sequence \( (s_n) \subset E \) which does not converge in \( E \). For \( k \in \mathbb{N} \), find \( n_k \) such that \( d(s_m, s_\ell) < 2^{-k} \) for \( m, \ell \geq n_k \). Construct open sets \( U_k = \{s \in S : d(s, s_{n_k}) > 2^{-k}\} \), which form an open cover for \( E \) with no finite subcover. Contradiction.
\end{proof}

\begin{corollary}
Any compact set is closed.
\end{corollary}

\begin{proposition}[Compactness Criterion]
\( E \subset S \) is compact if and only if it is complete and totally bounded.
\end{proposition}

\subsection*{Total Boundedness}
\begin{definition}[Total Boundedness]
A set \( E \subset S \) is called totally bounded if \( \forall \varepsilon > 0 \), there exist \( s_1, \ldots, s_n \in S \) such that:
\[
E \subset \bigcup_{i=1}^n B_\varepsilon^o(s_i).
\]
\end{definition}

\begin{proposition}
A set is totally bounded if and only if any sequence in the set has a Cauchy subsequence.
\end{proposition}

\subsection*{Characterization of Compactness}
\begin{theorem}
For a subset \( E \) of a metric space, the following are equivalent:
\begin{enumerate}
    \item \( E \) is compact.
    \item \( E \) is complete and totally bounded.
    \item Any sequence in \( E \) has a subsequence with a limit in \( E \).
\end{enumerate}
\end{theorem}

\begin{example}
The space \( \mathbb{N} \) with the discrete metric is complete and bounded but not compact.
\end{example}

\subsection*{Heine-Borel Theorem}
\begin{theorem}[Heine-Borel Theorem]
A subset of \( \mathbb{R}^n \) is compact if and only if it is closed and bounded.
\end{theorem}

\begin{example}[Cantor Set]
The Cantor set \( C \), constructed as:
\[
F_0 = [0, 1], \quad F_1 = [0, 1/3] \cup [2/3, 1], \quad F_2 = \cdots,
\]
is compact, closed, and totally bounded but has no interior.
\end{example}
\section*{A Note on Compactness}

\subsection*{Total Boundedness}
\begin{definition}[Total Boundedness (Definition 1.1)]
A set \( S \subset E \) is called totally bounded if for every \( \varepsilon > 0 \), there exist \( p_1, \ldots, p_n \in E \) such that:
\[
S \subset \bigcup_{i=1}^n B_\varepsilon^o(p_i).
\]
\end{definition}

\begin{proposition}[Intrinsic Nature of Total Boundedness (Proposition 1.2)]
A set \( S \subset E \) is totally bounded if and only if for every \( \varepsilon > 0 \), there exist \( q_1, \ldots, q_m \in S \) such that:
\[
S \subset \bigcup_{j=1}^m B_\varepsilon^o(q_j).
\]
\end{proposition}

\begin{proposition}[Total Boundedness and Cauchy Subsequences (Proposition 1.3)]
A set \( S \) is totally bounded if and only if any sequence in \( S \) has a Cauchy subsequence.
\end{proposition}

\begin{corollary}[Characterization of Compactness (Corollary 1.4)]
A set \( S \) is totally bounded and complete if and only if any sequence in \( S \) has a subsequence converging to a limit in \( S \).
\end{corollary}

\subsection*{Compactness}
\begin{theorem}[Characterization of Compactness (Theorem 2.1)]
For a subset \( S \subset E \), the following are equivalent:
\begin{enumerate}
    \item \( S \) is compact.
    \item Any sequence in \( S \) has a convergent subsequence.
    \item \( S \) is complete and totally bounded.
\end{enumerate}
\end{theorem}

\subsection*{Compactness in \( \mathbb{R}^n \)}
\begin{theorem}[Heine-Borel Theorem (Theorem 3.1)]
A set \( S \subset \mathbb{R}^n \) is compact if and only if it is closed and bounded.
\end{theorem}

\begin{lemma}[Total Boundedness in \( \mathbb{R}^n \) (Lemma 3.3)]
A set \( S \subset \mathbb{R}^n \) is bounded if and only if it is totally bounded.
\end{lemma}

\begin{theorem}[Bolzano-Weierstrass Theorem (Theorem 3.4)]
Every bounded sequence in \( \mathbb{R}^n \) has a convergent subsequence.
\end{theorem}
\section*{Compactness and Series (Sections 13-14)}

\subsection*{Total Boundedness}
\begin{definition}[Total Boundedness]
A set \( E \subset S \) is totally bounded if:
\[
\forall \varepsilon > 0, \exists s_1, \ldots, s_n \in S \text{ such that } E \subset \bigcup_{i=1}^n B_\varepsilon^o(s_i).
\]
\end{definition}

\begin{proposition}
A set \( E \) is totally bounded if and only if any sequence in \( E \) has a Cauchy subsequence.
\end{proposition}

\begin{proof}
For \( (s_i) \subset E \), construct \( \{s_{i_k}\} \) with \( s_{i_k} \in B_{2^{-k}}^o(x_{kj_k}) \). Using the triangle inequality, show \( (s_{i_k}) \) is Cauchy.
\end{proof}

\subsection*{Characterization of Compactness}
\begin{theorem}
For a subset \( E \) of a metric space, the following are equivalent:
\begin{enumerate}
    \item \( E \) is compact.
    \item \( E \) is complete and totally bounded.
    \item Any sequence in \( E \) has a subsequence with a limit in \( E \).
\end{enumerate}
\end{theorem}

\begin{proof}
\leavevmode
\begin{itemize}
    \item \( (1) \implies (2) \): Shown in the last lecture.
    \item \( (2) \implies (3) \): Total boundedness guarantees a Cauchy subsequence, and completeness ensures convergence.
    \item \( (3) \implies (1) \): Contraposition: if \( E \) is not compact, construct an open cover with no finite subcover.
\end{itemize}
\end{proof}

\subsection*{Compact Subsets of \( \mathbb{R}^n \)}
\begin{theorem}[Heine-Borel]
A subset of \( \mathbb{R}^n \) is compact if and only if it is closed and bounded.
\end{theorem}

\begin{example}[Non-Compact Set]
In \( \mathbb{N} \) with the discrete metric, \( \mathbb{N} \) is closed and bounded but not compact.
\end{example}

\subsection*{Series}
\begin{definition}[Series and Convergence]
The \( n \)-th partial sum of a series \( \sum_{j=k_0}^\infty a_j \) is:
\[
s_n = \sum_{j=k_0}^n a_j.
\]
The series converges if \( \lim_{n \to \infty} s_n \) exists, diverges otherwise.
\end{definition}

\begin{example}[Geometric Series]
\[
\sum_{j=0}^\infty r^j =
\begin{cases} 
\frac{1}{1-r}, & |r| < 1, \\
\infty, & r \geq 1.
\end{cases}
\]
\end{example}

\subsection*{Cauchy Criterion for Convergence}
\begin{definition}[Cauchy Criterion]
A series \( \sum_j a_j \) satisfies the Cauchy criterion if:
\[
\forall \varepsilon > 0, \exists N \text{ such that } \left| \sum_{j=m}^n a_j \right| < \varepsilon \text{ for } n \geq m > N.
\]
\end{definition}

\begin{theorem}
A series converges if and only if it satisfies the Cauchy criterion.
\end{theorem}

\subsection*{Comparison Test for Convergence}
\begin{theorem}[Comparison Test]
\leavevmode
\begin{enumerate}
    \item If \( |b_n| \leq a_n \) and \( \sum a_n \) converges, then \( \sum b_n \) converges.
    \item If \( 0 \leq a_n \leq b_n \) and \( \sum b_n = \infty \), then \( \sum a_n = \infty \).
\end{enumerate}
\end{theorem}
\section*{Series and Decimal Expansions (Sections 14 and 16)}

\subsection*{Series}
\begin{definition}[Partial Sums and Convergence of Series]
The \( n \)-th partial sum of a series \( \sum_{j=k_0}^\infty a_j \) is:
\[
s_n = \sum_{j=k_0}^n a_j.
\]
The series \( \sum_{j=k_0}^\infty a_j \) converges if \( \lim_{n \to \infty} s_n \) exists. Otherwise, it diverges.
\end{definition}

\begin{example}[Geometric Series]
\[
\sum_{j=0}^\infty r^j =
\begin{cases}
\frac{1}{1-r}, & |r| < 1, \\
\text{diverges}, & r \geq 1 \text{ or } r \leq -1.
\end{cases}
\]
\end{example}

\subsection*{Cauchy Criterion for Convergence}
\begin{definition}[Cauchy Criterion (Definition 14.3)]
A series \( \sum_j a_j \) satisfies the Cauchy Criterion if:
\[
\forall \varepsilon > 0, \exists N \text{ such that } \left|\sum_{j=m}^n a_j\right| < \varepsilon \text{ for } n \geq m > N.
\]
\end{definition}

\begin{theorem}[Cauchy Criterion (Theorem 14.4)]
A series converges if and only if it satisfies the Cauchy Criterion.
\end{theorem}

\subsection*{Properties of Convergence}
\begin{corollary}[Necessary Condition for Convergence (Corollary 14.5)]
If \( \sum_j a_j \) converges, then \( \lim_{n \to \infty} a_n = 0 \).
\end{corollary}

\begin{example}
If \( a_n = \frac{1}{n} \), then \( \lim_{n \to \infty} a_n = 0 \), but \( \sum_{n=1}^\infty \frac{1}{n} \) diverges.
\end{example}

\subsection*{Tests for Convergence}
\begin{theorem}[Comparison Test (Theorem 14.6)]
\leavevmode
\begin{enumerate}
    \item If \( 0 \leq |b_n| \leq a_n \) and \( \sum a_n \) converges, then \( \sum b_n \) converges.
    \item If \( 0 \leq a_n \leq b_n \) and \( \sum b_n = \infty \), then \( \sum a_n = \infty \).
\end{enumerate}
\end{theorem}

\begin{theorem}[Root Test (Theorem 14.9)]
For a series \( \sum_n a_n \), let \( \alpha = \limsup_{n \to \infty} |a_n|^{1/n} \). Then:
\begin{enumerate}
    \item The series converges absolutely if \( \alpha < 1 \).
    \item The series diverges if \( \alpha > 1 \).
    \item If \( \alpha = 1 \), the test gives no information.
\end{enumerate}
\end{theorem}

\begin{theorem}[Ratio Test (Theorem 14.8)]
For a series \( \sum_n a_n \) of nonzero terms:
\begin{enumerate}
    \item The series converges absolutely if \( \limsup_{n \to \infty} \left|\frac{a_{n+1}}{a_n}\right| < 1 \).
    \item The series diverges if \( \liminf_{n \to \infty} \left|\frac{a_{n+1}}{a_n}\right| > 1 \).
    \item If \( \liminf_{n \to \infty} \left|\frac{a_{n+1}}{a_n}\right| \leq 1 \leq \limsup_{n \to \infty} \left|\frac{a_{n+1}}{a_n}\right| \), the test gives no information.
\end{enumerate}
\end{theorem}

\subsection*{Decimal Expansions}
\begin{theorem}[Decimal Expansions (Theorem 16.2)]
Any real number \( x \geq 0 \) has at least one decimal expansion:
\[
x = K.d_1d_2d_3\ldots = K + \sum_{j=1}^\infty \frac{d_j}{10^j},
\]
where \( K \in \mathbb{Z} \) and \( d_j \in \{0, 1, \ldots, 9\} \).
\end{theorem}

\begin{theorem}[Uniqueness of Decimal Expansions (Theorem 16.3)]
Any \( x \geq 0 \) has either exactly one decimal expansion or exactly two, one ending in \( \ldots d000\ldots \) and the other in \( \ldots [d-1]999\ldots \).
\end{theorem}

\begin{theorem}[Repeating Decimals (Theorem 16.5)]
A real number \( x \) is rational if and only if its decimal expansion is repeating.
\end{theorem}
\section*{Series and Decimal Expansions (Sections 17 and 21)}

\subsection*{Root and Ratio Tests for Series Convergence}
\begin{theorem}[Root Test]
For a series \( \sum_{n} a_n \), let \( \alpha = \limsup_{n \to \infty} |a_n|^{1/n} \). Then:
\begin{enumerate}
    \item The series converges absolutely if \( \alpha < 1 \).
    \item The series diverges if \( \alpha > 1 \).
    \item If \( \alpha = 1 \), the test gives no information.
\end{enumerate}
\end{theorem}

\begin{theorem}[Ratio Test (Theorem 14.8)]
For a series \( \sum_{n} a_n \) of nonzero terms:
\begin{enumerate}
    \item The series converges absolutely if \( \limsup_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| < 1 \).
    \item The series diverges if \( \liminf_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| > 1 \).
    \item If \( \liminf_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| \leq 1 \leq \limsup_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| \), the test gives no information.
\end{enumerate}
\end{theorem}

\begin{example}
\leavevmode
\begin{itemize}
    \item Consider \( \sum_{k=1}^\infty \frac{k^4}{2^k} \). Using the Root Test:
    \[
    a_k^{1/k} = \left( k^{1/k} \right)^4 \frac{1}{2}, \quad \lim_{k \to \infty} a_k^{1/k} = \frac{1}{2} < 1,
    \]
    so the series converges.
    \item The \( p \)-series \( \sum_{k=1}^\infty \frac{1}{k^p} \) converges if and only if \( p > 1 \). The Root and Ratio Tests are inconclusive for this series.
\end{itemize}
\end{example}

\subsection*{Decimal Expansions}
\begin{definition}[Decimal Expansion]
For \( x \in [0, \infty) \), the decimal expansion of \( x \) is:
\[
x = K .d_1 d_2 d_3 \ldots = K + \sum_{j=1}^\infty \frac{d_j}{10^j},
\]
where \( K \in \{0, 1, 2, \ldots\} \) and \( d_1, d_2, \ldots \in \{0, 1, \ldots, 9\} \).
\end{definition}

\begin{theorem}[Existence of Decimal Expansions (Theorem 16.2)]
Any real number \( x \geq 0 \) has at least one decimal expansion.
\end{theorem}

\begin{theorem}[Uniqueness of Decimal Expansions (Theorem 16.3)]
Any \( x \geq 0 \) has either exactly one decimal expansion or exactly two:
\begin{itemize}
    \item One ending in \( \ldots d000 \ldots \), where \( d \in \{1, \ldots, 9\} \),
    \item Another ending in \( \ldots (d-1)999 \ldots \).
\end{itemize}
For example, \( \frac{1}{2} = 0.5000\ldots = 0.4999\ldots \).
\end{theorem}

\subsection*{Repeating Decimal Expansions}
\begin{definition}[Repeating Decimals (Definition 16.4)]
A repeating decimal expansion is one of the form:
\[
K .d_1 \ldots d_\ell d_{\ell+1} \ldots d_{\ell+r} = K .d_1 \ldots d_\ell \overline{d_{\ell+1} \ldots d_{\ell+r}},
\]
where the sequence \( d_{\ell+1} \ldots d_{\ell+r} \) repeats.
\end{definition}

\begin{theorem}[Repeating Decimals and Rational Numbers (Theorem 16.5)]
A real number \( x \) is rational if and only if its decimal expansion is repeating.
\end{theorem}

\begin{proof}
\leavevmode
\begin{itemize}
    \item (\( x \) is rational \( \implies \) repeating): Follows from performing long division.
    \item (Repeating \( \implies x \) is rational): Suppose \( x = K .d_1 \ldots d_\ell \overline{d_{\ell+1} \ldots d_{\ell+r}} \). Then:
    \[
    x = K + \sum_{j=1}^\ell \frac{d_j}{10^j} + 10^{-\ell} \left( \frac{z}{1 - 10^{-r}} \right),
    \]
    where \( z = \sum_{j=1}^r d_{\ell+j} 10^{-j} \in \mathbb{Q} \), so \( x \in \mathbb{Q} \).
\end{itemize}
\end{proof}
\section*{Continuity in Metric Spaces (Sections 17 and 21)}

\subsection*{Definition of Continuity}
\begin{definition}[Continuity (Definition 21.1)]
Suppose \( (S, d) \) and \( (S^*, d^*) \) are metric spaces. The function \( f : \text{dom}(f) \to S^* \) (with \( \text{dom}(f) \subset S \)) is continuous at \( x \in \text{dom}(f) \) if:
\[
\forall \varepsilon > 0, \exists \delta > 0 \text{ such that } d^*(f(x), f(y)) < \varepsilon \text{ whenever } d(x, y) < \delta.
\]
\( f \) is called continuous on \( E \subset S \) if it is continuous at every \( x \in E \).
\end{definition}

\begin{theorem}[Sequential Criterion for Continuity (Theorem 17.1 + 17.2)]
\( f : S \to S^* \) is continuous at \( x \in S \) if and only if \( f(x_n) \to f(x) \) whenever \( x_n \to x \).
\end{theorem}

\subsection*{Examples of Continuity and Discontinuity}
\begin{example}[Discontinuous Everywhere]
The Dirichlet function \( f : \mathbb{R} \to \mathbb{R} \), defined as:
\[
f(x) =
\begin{cases}
1 & x \in \mathbb{Q}, \\
0 & x \notin \mathbb{Q},
\end{cases}
\]
is discontinuous at every \( x \in \mathbb{R} \). This is because for any \( x \in \mathbb{R} \), one can find sequences \( (x_n) \subset \mathbb{Q} \) and \( (y_n) \not\subset \mathbb{Q} \) such that \( x_n, y_n \to x \), but \( f(x_n) \to 1 \) and \( f(y_n) \to 0 \), which do not match \( f(x) \).
\end{example}

\begin{example}[Continuous at a Single Point]
The modified Dirichlet function \( g : \mathbb{R} \to \mathbb{R} \), defined as:
\[
g(x) =
\begin{cases}
x & x \in \mathbb{Q}, \\
0 & x \notin \mathbb{Q},
\end{cases}
\]
is continuous only at \( x = 0 \). At other points, similar reasoning as the Dirichlet function applies.
\end{example}

\begin{example}[Continuous on \( \mathbb{R} \setminus \mathbb{Q} \)]
The Thomae function \( h : \mathbb{R} \to \mathbb{R} \), defined as:
\[
h(x) =
\begin{cases}
\frac{1}{b} & x = \frac{a}{b}, \, \gcd(a, b) = 1, \, b > 0, \, x \neq 0, \\
0 & x \notin \mathbb{Q},
\end{cases}
\]
is continuous at \( x \notin \mathbb{Q} \) and discontinuous at \( x \in \mathbb{Q} \).
\end{example}

\subsection*{Operations on Continuous Functions}
\begin{theorem}
Suppose \( f, g \) are continuous at \( x_0 \) in a metric space \( (S, d) \). Then the following functions are also continuous at \( x_0 \):
\begin{itemize}
    \item \( |f| \),
    \item \( kf \) (\( k \in \mathbb{R} \)),
    \item \( f + g \),
    \item \( f \cdot g \),
    \item \( f / g \) (if \( g(x_0) \neq 0 \)).
\end{itemize}
\end{theorem}

\begin{proposition}
If \( f, g \) are continuous at \( x_0 \), then \( \max(f, g) \) and \( \min(f, g) \) are continuous at \( x_0 \).
\end{proposition}

\subsection*{Composition of Continuous Functions}
\begin{theorem}
Suppose \( (S_1, d_1), (S_2, d_2), (S_3, d_3) \) are metric spaces, and \( f : \text{dom}(f) \to S_2 \), \( g : \text{dom}(g) \to S_3 \) are functions such that \( f \) is continuous at \( x_0 \), \( g \) is continuous at \( f(x_0) \), and \( x_0 \in \text{dom}(f) \). Then \( g \circ f \) is continuous at \( x_0 \).
\end{theorem}

\subsection*{Characterization of Continuity}
\begin{theorem}[Characterization of Continuity (Theorem 21.3)]
Suppose \( (S, d) \) and \( (S^*, d^*) \) are metric spaces. \( f : S \to S^* \) is continuous if and only if \( f^{-1}(U) \) is open for every open \( U \subset S^* \), where:
\[
f^{-1}(U) = \{s \in S : f(s) \in U\}.
\]
\end{theorem}

\begin{lemma}
\( f \) is continuous at \( s_0 \in S \) if for any open set \( U \) containing \( f(s_0) \), there exists an open set \( V \) containing \( s_0 \) such that \( f(V) \subset U \).
\end{lemma}
\section*{Continuity and the Intermediate Value Theorem (Sections 18 and 21)}

\subsection*{Another Characterization of Continuity}
\begin{theorem}[Theorem 21.3]
Suppose \( (S, d) \) and \( (S^*, d^*) \) are metric spaces. A function \( f : S \to S^* \) is continuous if and only if \( f^{-1}(U) \) is open for every open \( U \subset S^* \). Here:
\[
f^{-1}(U) = \{s \in S : f(s) \in U\}.
\]
\end{theorem}

\begin{lemma}[Exercise 21.2]
\( f \) is continuous at \( s_0 \in S \) if and only if for any open set \( U \ni f(s_0) \), there exists an open set \( V \ni s_0 \) such that \( f(V) \subset U \).
\end{lemma}

\begin{corollary}[Exercise 21.4]
Suppose \( (S, d) \) is a metric space. A function \( f : S \to \mathbb{R} \) is continuous if and only if \( f^{-1}((a, b)) \) is open whenever \( a < b \).
\end{corollary}

\subsection*{Continuous Image of a Compact Set}
\begin{theorem}
If \( f : S \to S^* \) is continuous, and \( E \subset S \) is compact, then \( f(E) \subset S^* \) is compact.
\end{theorem}

\begin{corollary}
If \( f : S \to \mathbb{R} \) is continuous, and \( E \subset S \) is compact, then \( f(E) \) is bounded. Moreover, \( f \) attains its maximum and minimum values, i.e., there exist \( x, y \in E \) such that:
\[
f(x) = \sup_{e \in E} f(e), \quad f(y) = \inf_{e \in E} f(e).
\]
\end{corollary}

\subsection*{Intermediate Value Theorem (IVT)}
\begin{theorem}[Theorem 18.2]
Suppose \( I \subset \mathbb{R} \) is an interval, and \( f : I \to \mathbb{R} \) is continuous. Then \( f \) has the Intermediate Value Property (IVP) on \( I \): if \( a, b \in I \) with \( a < b \), and \( y \) lies between \( f(a) \) and \( f(b) \), then there exists \( x \in (a, b) \) such that \( f(x) = y \).
\end{theorem}

\begin{corollary}
If \( I \) is an interval, and \( f : I \to \mathbb{R} \) has the IVP, then \( f(I) \) is either an interval or a single point.
\end{corollary}

\subsection*{Applications of IVT}
\begin{proposition}[Roots of Polynomials]
Any polynomial of odd degree has at least one real root.
\end{proposition}

\begin{proposition}[Existence of Fixed Points]
Any continuous function \( f : [0, 1] \to [0, 1] \) has a fixed point, i.e., \( x \in [0, 1] \) such that \( f(x) = x \).
\end{proposition}

\begin{proposition}[Existence of \( m \)-th Roots]
For any \( m \in \mathbb{N} \) and \( y > 0 \), there exists \( x > 0 \) such that \( x^m = y \).
\end{proposition}

\subsection*{Continuity of Inverse Functions}
\begin{theorem}[Theorem 18.4]
Suppose \( I \subset \mathbb{R} \) is an interval, and \( f : I \to \mathbb{R} \) is strictly increasing and continuous. Then \( J = f(I) \) is an interval, and \( f^{-1} : J \to I \) is strictly increasing and continuous.
\end{theorem}

\begin{corollary}
The function \( x \mapsto x^{1/m} \), taking \( [0, \infty) \) to itself, is continuous.
\end{corollary}
\section*{Continuity and Compactness (Section 18)}

\subsection*{Continuous Image of a Compact Set}
\begin{theorem}[Theorem 21.4(i)]
Suppose \( f : S \to S^* \) is continuous, where \( (S, d) \) and \( (S^*, d^*) \) are metric spaces, and \( E \subset S \) is compact. Then \( f(E) \subset S^* \) is compact.
\end{theorem}

\begin{proof}
Let \( (U_i)_{i \in I} \) be an open cover for \( f(E) \). Define \( V_i = f^{-1}(U_i) \), which are open sets forming a cover for \( E \). By compactness of \( E \), there exist \( i_1, \ldots, i_n \in I \) such that \( E \subset \bigcup_{k=1}^n V_{i_k} \). It follows that \( f(E) \subset \bigcup_{k=1}^n U_{i_k} \), proving compactness of \( f(E) \).
\end{proof}

\subsection*{Maximum and Minimum of Continuous Functions}
\begin{corollary}[Similar to 18.1]
If \( f : S \to \mathbb{R} \) is continuous and \( E \subset S \) is compact, then \( f(E) \) is bounded. Moreover, \( f \) attains its maximum and minimum values, i.e., there exist \( x, y \in E \) such that:
\[
f(x) = \sup_{e \in E} f(e), \quad f(y) = \inf_{e \in E} f(e).
\]
\end{corollary}

\subsection*{Intermediate Value Property}
\begin{definition}[IVP]
Suppose \( I \subset \mathbb{R} \) is an interval, and \( f : I \to \mathbb{R} \) is a function. \( f \) has the Intermediate Value Property (IVP) on \( I \) if for any \( a, b \in I \) with \( a < b \), and any \( y \) between \( f(a) \) and \( f(b) \), there exists \( x \in (a, b) \) such that \( f(x) = y \).
\end{definition}

\begin{theorem}[Theorem 18.2]
Any continuous function has the IVP.
\end{theorem}

\subsection*{Applications of IVP}
\begin{corollary}[18.3]
If \( I \) is an interval, and \( f : I \to \mathbb{R} \) has the IVP, then \( f(I) \) is either an interval or a single point.
\end{corollary}

\begin{proposition}[Roots of Polynomials]
Any polynomial of odd degree has at least one real root.
\end{proposition}

\begin{proposition}[Existence of Fixed Points]
Any continuous function \( f : [0, 1] \to [0, 1] \) has a fixed point, i.e., a point \( x \in [0, 1] \) such that \( f(x) = x \).
\end{proposition}

\begin{proposition}[Existence of \( m \)-th Root]
For any \( m \in \mathbb{N} \) and \( y > 0 \), there exists \( x > 0 \) such that \( x^m = y \).
\end{proposition}

\subsection*{Continuity of Inverse Functions}
\begin{theorem}[Theorem 18.4]
Suppose \( I \subset \mathbb{R} \) is an interval, and \( f : I \to \mathbb{R} \) is strictly increasing and continuous. Then \( f(I) \) is an interval, and \( f^{-1} : f(I) \to I \) is strictly increasing and continuous.
\end{theorem}

\begin{corollary}
The function \( x \mapsto x^{1/m} \), taking \( [0, \infty) \) to itself, is continuous.
\end{corollary}

\subsection*{Monotonicity of Injective Functions}
\begin{theorem}[Theorem 18.6]
Suppose \( f : I \to \mathbb{R} \) is a continuous one-to-one function on an interval \( I \). Then \( f \) is strictly monotone.
\end{theorem}

\begin{proof}[Sketch]
For \( a, b \in I \) with \( a < b \), if \( f(a) < f(b) \), then \( f \) is strictly increasing. Otherwise, by the IVP, there would exist \( x \in (a, b) \) such that \( f(x) = f(a) \), contradicting injectivity.
\end{proof}
\section*{Uniform Continuity and Lipschitz Functions (Sections 18-19)}

\subsection*{Uniform Continuity}
\begin{definition}[Uniform Continuity (Definition 21.1)]
Suppose \( (S, d) \) and \( (S^*, d^*) \) are metric spaces. A function \( f : S \to S^* \) is uniformly continuous on \( E \subset S \) if:
\[
\forall \varepsilon > 0, \exists \delta > 0 \text{ such that } d^*(f(x), f(y)) < \varepsilon \text{ whenever } d(x, y) < \delta.
\]
Here, \( \delta \) depends only on \( \varepsilon \) and not on the specific point \( x \).
\end{definition}

\begin{theorem}[Sequential Criterion for Uniform Continuity (Theorem 19.4)]
If \( f : S \to S^* \) is uniformly continuous, then \( f \) maps Cauchy sequences in \( S \) to Cauchy sequences in \( S^* \).
\end{theorem}

\begin{example}[Non-Uniformly Continuous Function]
The function \( f(x) = \frac{1}{x} \) is not uniformly continuous on \( (0, \infty) \), as the Cauchy sequence \( x_n = \frac{1}{n} \) is mapped to \( f(x_n) = n \), which is not Cauchy.
\end{example}

\subsection*{Lipschitz Functions}
\begin{definition}[Lipschitz Continuity]
A function \( f : S \to S^* \) is Lipschitz if there exists \( K > 0 \) (called the Lipschitz constant) such that:
\[
d^*(f(s), f(t)) \leq K \cdot d(s, t), \quad \forall s, t \in S.
\]
\end{definition}

\begin{proposition}
Any Lipschitz function is uniformly continuous.
\end{proposition}

\begin{proof}
Let \( \varepsilon > 0 \) and set \( \delta = \frac{\varepsilon}{K} \). Then if \( d(s, t) < \delta \), it follows that:
\[
d^*(f(s), f(t)) \leq K \cdot d(s, t) < K \cdot \frac{\varepsilon}{K} = \varepsilon.
\]
\end{proof}

\begin{example}
For \( a > 0 \), \( f(x) = \frac{1}{x} \) is Lipschitz (hence uniformly continuous) on \( [a, \infty) \).
\end{example}

\begin{example}[Uniformly Continuous but Not Lipschitz]
The function \( f(x) = \sqrt{x} \) is uniformly continuous on \( [0, \infty) \) but not Lipschitz.
\end{example}

\subsection*{Uniform Continuity on Compact Sets}
\begin{theorem}[Uniform Continuity on Compact Sets (Theorem 21.4(ii))]
If \( f : S \to S^* \) is continuous, and \( E \subset S \) is compact, then \( f \) is uniformly continuous on \( E \).
\end{theorem}

\begin{proof}[Sketch of Proof]
Assume \( f \) is not uniformly continuous. Then \( \exists \varepsilon > 0 \) and sequences \( (x_n), (y_n) \subset E \) such that \( d(x_n, y_n) \to 0 \) but \( d^*(f(x_n), f(y_n)) \geq \varepsilon \). By compactness, \( (x_n) \) has a subsequence \( (x_{n_k}) \) converging to some \( x \in E \), and \( (y_{n_k}) \) also converges to \( x \). Continuity of \( f \) implies \( d^*(f(x_{n_k}), f(y_{n_k})) \to 0 \), contradicting \( d^*(f(x_{n_k}), f(y_{n_k})) \geq \varepsilon \).
\end{proof}
\section*{Uniform Continuity and Connectedness (Section 22)}

\subsection*{Uniform Continuity on Compact Sets}
\begin{theorem}[Theorem 21.4(ii)]
Suppose \( (S, d) \) and \( (S^*, d^*) \) are metric spaces, and \( f : S \to S^* \) is continuous. If \( E \subset S \) is compact, then \( f|_E \) is uniformly continuous.
\end{theorem}

\begin{proof}
For \( \varepsilon > 0 \), find \( \delta > 0 \) such that \( d^*(f(s), f(t)) < \varepsilon \) whenever \( d(s, t) < \delta \). For \( s \in S \), find \( \delta_s > 0 \) such that \( d^*(f(s), f(t)) < \varepsilon/2 \) whenever \( d(s, t) < \delta_s \). Since \( E \) is compact:
\[
E \subset \bigcup_{s \in E} B_{\delta_s/2}^o(s),
\]
there exist \( s_1, \ldots, s_n \) such that \( E \subset \bigcup_{k=1}^n B_{\delta_{s_k}/2}^o(s_k) \). Define \( \delta = \frac{1}{2} \min_{1 \leq k \leq n} \delta_{s_k} \). For \( s, t \in E \) with \( d(s, t) < \delta \), choose \( s_k \) such that \( s \in B_{\delta_{s_k}/2}^o(s_k) \). Then:
\[
d(t, s_k) \leq d(t, s) + d(s, s_k) < \delta + \frac{\delta_{s_k}}{2} \leq \delta_{s_k},
\]
implying \( d^*(f(s), f(t)) \leq d^*(f(s), f(s_k)) + d^*(f(t), f(s_k)) < \varepsilon \).
\end{proof}

\subsection*{Extension of Uniformly Continuous Functions}
\begin{theorem}
Suppose \( E \subset S \) is compact, \( f : E \to S^* \), and \( S^* \) is complete. Then \( f \) is uniformly continuous if and only if it extends to a continuous \( \tilde{f} : \overline{E} \to S^* \).
\end{theorem}

\begin{proof}[Sketch]
If \( f \) is uniformly continuous, define \( \tilde{f}(x) = \lim_{n \to \infty} f(x_n) \) for \( x \in \overline{E} \), where \( x_n \subset E \) and \( x_n \to x \). The limit exists by completeness and does not depend on the sequence. Continuity of \( \tilde{f} \) follows from the uniform continuity of \( f \).
\end{proof}

\subsection*{Connectedness}
\begin{definition}[Connected and Disconnected Sets]
Suppose \( (S, d) \) is a metric space. A set \( E \subset S \) is disconnected if there exist open sets \( U_1, U_2 \subset S \) such that:
\begin{enumerate}
    \item \( E \subset U_1 \cup U_2 \),
    \item \( (E \cap U_1) \cap (E \cap U_2) = \emptyset \),
    \item \( E \cap U_1 \neq \emptyset \) and \( E \cap U_2 \neq \emptyset \).
\end{enumerate}
A set \( E \) is connected if it is not disconnected.
\end{definition}

\begin{proposition}
\( E \) is disconnected if and only if there exist \( A, B \subset E \) such that:
\[
E = A \cup B, \quad A \neq \emptyset, \quad B \neq \emptyset, \quad A \cap \overline{B} = \emptyset, \quad \overline{A} \cap B = \emptyset.
\]
\end{proposition}

\subsection*{Connectedness of Intervals}
\begin{proposition}
Any interval \( I \subset \mathbb{R} \) is connected.
\end{proposition}

\begin{proof}
Suppose, for contradiction, that \( I = A \cup B \), where \( A, B \neq \emptyset \), \( \overline{A} \cap B = \emptyset \), and \( A \cap \overline{B} = \emptyset \). Choose \( a \in A \), \( b \in B \), with \( a < b \). Define:
\[
c = \sup \{x \in A : x < b\}.
\]
Then \( c \in I \) and \( c < b \). If \( c \in A \), there exists \( \sigma > 0 \) such that \( (c - \sigma, c + \sigma) \subset A \), contradicting the definition of \( c \). If \( c \in B \), there exists \( \sigma > 0 \) such that \( (c - \sigma, c + \sigma) \subset B \), contradicting \( c = \sup \{x \in A : x < b\} \).
\end{proof}
\section*{Connectedness and Path Connectedness (Section 22)}

\subsection*{Connectedness}
\begin{definition}[Connected Set (Definition 22.1)]
Suppose \( (S, d) \) is a metric space. A set \( E \subset S \) is called disconnected if there exist open sets \( U_1, U_2 \subset S \) such that:
\begin{enumerate}
    \item \( E \subset U_1 \cup U_2 \),
    \item \( (E \cap U_1) \cap (E \cap U_2) = \emptyset \),
    \item \( E \cap U_1 \neq \emptyset \) and \( E \cap U_2 \neq \emptyset \).
\end{enumerate}
A set \( E \) is connected if it is not disconnected.
\end{definition}

\begin{proposition}
An open set \( E \) is disconnected if and only if \( E = E_1 \cup E_2 \), where \( E_1, E_2 \) are disjoint, non-empty, open subsets.
\end{proposition}

\begin{proposition}[Equivalent Characterization of Connectedness]
A set \( E \) is disconnected if and only if there exist \( A, B \subset E \) such that:
\[
E = A \cup B, \quad A \neq \emptyset, \quad B \neq \emptyset, \quad A \cap \overline{B} = \emptyset, \quad \overline{A} \cap B = \emptyset.
\]
\end{proposition}

\subsection*{Continuous Images of Connected Sets}
\begin{theorem}[Theorem 22.2]
Suppose \( (S, d) \) and \( (S^*, d^*) \) are metric spaces. If \( E \subset S \) is connected and \( f : S \to S^* \) is continuous, then \( f(E) \) is connected.
\end{theorem}

\begin{proof}[Sketch of Contrapositive Proof]
If \( f(E) \subset S^* \) is disconnected, write \( f(E) = C \cup D \), where \( C, D \) are disjoint, non-empty, closed subsets. Define \( A = f^{-1}(C) \cap E \), \( B = f^{-1}(D) \cap E \). Then \( E = A \cup B \), \( A \cap \overline{B} = \emptyset \), and \( \overline{A} \cap B = \emptyset \), so \( E \) is disconnected.
\end{proof}

\subsection*{Path Connectedness}
\begin{definition}[Path Connectedness (Definition 22.4)]
A set \( E \subset S \) is path connected if for all \( a, b \in E \), there exists a continuous function \( \gamma : [0, 1] \to E \) such that \( \gamma(0) = a \) and \( \gamma(1) = b \).
\end{definition}

\begin{theorem}[Path Connected Sets Are Connected (Theorem 22.5)]
Every path connected set is connected.
\end{theorem}

\begin{proof}
If \( E \) is disconnected, then there exist open sets \( U_1, U_2 \) such that \( E \subset U_1 \cup U_2 \), \( E \cap U_1 \neq \emptyset \), \( E \cap U_2 \neq \emptyset \), and \( (E \cap U_1) \cap (E \cap U_2) = \emptyset \). Let \( a \in E \cap U_1 \), \( b \in E \cap U_2 \). A path \( \gamma : [0, 1] \to E \) with \( \gamma(0) = a \), \( \gamma(1) = b \) would imply \( \gamma([0, 1]) \) is connected, contradicting the disconnectedness of \( E \).
\end{proof}

\subsection*{Connected but Not Path Connected Sets}
\begin{example}
Consider \( E \subset \mathbb{R}^2 \), where:
\[
E_1 = \{(0, y) : y \in (0, 1]\}, \quad
E_2 = \{(x, 0) : x \in (0, 1]\} \cup \bigcup_{n \in \mathbb{N}} \{(1/n, y) : y \in (0, 1]\}.
\]
Then \( E = E_1 \cup E_2 \) is connected but not path connected.
\end{example}

\subsection*{Convex Sets}
\begin{definition}[Convex Sets]
A set \( E \subset \mathbb{R}^n \) is convex if for all \( \vec{x}, \vec{y} \in E \) and \( t \in [0, 1] \), the point:
\[
\vec{z} = (1-t)\vec{x} + t\vec{y} \in E.
\]
\end{definition}

\begin{proposition}
Any convex set is path connected.
\end{proposition}

\subsection*{Path Connectedness of Graphs}
\begin{proposition}
The graph of a function \( f : I \to \mathbb{R} \), where \( I \subset \mathbb{R} \) is an interval, is path connected if and only if \( f \) is continuous.
\end{proposition}
\section*{Graphs of Functions and Path Connectedness (Sections 22-23-24)}

\subsection*{Graphs and Path Connectedness}
\begin{definition}[Graph of a Function]
The graph of a function \( f : I \to \mathbb{R} \) (where \( I \subset \mathbb{R} \) is an interval) is:
\[
G(f) = \{(x, f(x)) : x \in I\}.
\]
\end{definition}

\begin{proposition}[Example 4 from Section 22]
\( G(f) \) is path connected if and only if \( f \) is continuous on \( I \).
\end{proposition}

\begin{example}[Discontinuous \( f \) with Connected \( G(f) \)]
Exercise 22.4 describes a function \( f \) such that \( G(f) \) is connected but \( f \) is discontinuous.
\end{example}

\begin{proposition}[Multivariate Continuity]
The function \( f : S \to \mathbb{R}^n \), \( x \mapsto (f_1(x), \ldots, f_n(x)) \), is continuous if and only if each \( f_i : S \to \mathbb{R} \) is continuous for \( 1 \leq i \leq n \).
\end{proposition}

\begin{proof}[Sketch]
If \( f \) is continuous, then \( G(f) \) is path connected. For \( \vec{x} = (a, f(a)) \), \( \vec{y} = (b, f(b)) \in G(f) \), define a path:
\[
\gamma(t) = ((1-t)a + tb, f((1-t)a + tb)), \quad t \in [0, 1].
\]
If \( G(f) \) is path connected, continuity of \( f \) follows from the textbook proof.
\end{proof}

\subsection*{Power Series}
\begin{definition}[Power Series]
A power series is a series of the form:
\[
\sum_{n=0}^\infty a_n x^n,
\]
where \( x \) is a variable.
\end{definition}

\begin{theorem}[Radius of Convergence]
Let \( \beta = \limsup |a_n|^{1/n} \) and \( R = 1/\beta \). The series:
\[
\sum_{n=0}^\infty a_n x^n
\]
converges for \( |x| < R \), diverges for \( |x| > R \). \( R \) is called the radius of convergence.
\end{theorem}

\begin{remark}
If \( \lim |a_{n+1}/a_n| \) exists, it equals \( \beta \). The series may converge or diverge at \( \pm R \). The interval of convergence is one of:
\[
(-R, R), \, [-R, R), \, (-R, R], \, \text{or} \, [-R, R].
\]
\end{remark}

\subsection*{Examples of Power Series}
\begin{enumerate}
    \item \( \sum_{n=0}^\infty \frac{x^n}{n!} \): \( a_n = \frac{1}{n!} \), \( \beta = 0 \), \( R = \infty \). Interval: \( (-\infty, \infty) \).
    \[
    \sum_{n=0}^\infty \frac{x^n}{n!} = e^x.
    \]
    \item \( \sum_{n=0}^\infty x^n \): \( a_n = 1 \), \( \beta = 1 \), \( R = 1 \). Diverges for \( x = \pm 1 \). Interval: \( (-1, 1) \).
    \[
    \sum_{n=0}^\infty x^n = \frac{1}{1-x}.
    \]
    \item \( \sum_{n=0}^\infty \frac{x^n}{n+1} \): \( a_n = \frac{1}{n+1} \), \( \beta = 1 \), \( R = 1 \). Diverges at \( x = 1 \), converges for \( x \in [-1, 1) \).
    \[
    \sum_{n=0}^\infty \frac{x^n}{n+1} = \ln(1-x).
    \]
    \item \( \sum_{n=0}^\infty \frac{x^n}{(n+1)^2} \): \( a_n = \frac{1}{(n+1)^2} \), \( \beta = 1 \), \( R = 1 \). Converges for \( x \in [-1, 1] \).
    \item \( \sum_{n=0}^\infty n! x^n \): \( a_n = n! \), \( \beta = \infty \), \( R = 0 \). Diverges for all \( x \neq 0 \).
\end{enumerate}

\subsection*{Uniform Convergence}
\begin{definition}[Uniform Convergence (Definition 24.1-2)]
A sequence \( f_n \to f \) pointwise on \( S \) if:
\[
\forall x \in S, \, \forall \varepsilon > 0, \exists N \text{ such that } |f_n(x) - f(x)| < \varepsilon \text{ for } n \geq N.
\]
It converges uniformly if:
\[
\forall \varepsilon > 0, \exists N \text{ such that } \sup_{x \in S} |f_n(x) - f(x)| < \varepsilon \text{ for } n \geq N.
\]
\end{definition}

\begin{theorem}[Preservation of Continuity (Theorem 24.3)]
If \( f_n \to f \) uniformly on \( S \) and each \( f_n \) is continuous, then \( f \) is continuous.
\end{theorem}

\begin{example}
Consider \( f_n(x) = n^2 x^n (1-x) \) on \( [0, 1] \). It converges pointwise to \( f(x) = 0 \), but not uniformly.
\end{example}
\section*{Uniform Convergence and Series of Functions (Sections 24-25)}

\subsection*{Uniform Convergence}
\begin{definition}[Pointwise and Uniform Convergence (24.1-2)]
Suppose \( f, f_1, f_2, \ldots \) are functions \( S \to \mathbb{R} \).
\begin{itemize}
    \item \( f_n \to f \) pointwise on \( S \) if:
    \[
    \forall x \in S, \, \forall \varepsilon > 0, \, \exists N \in \mathbb{N} \text{ such that } |f_n(x) - f(x)| < \varepsilon \text{ for } n \geq N.
    \]
    \item \( f_n \to f \) uniformly on \( S \) if:
    \[
    \forall \varepsilon > 0, \, \exists N \in \mathbb{N} \text{ such that } |f_n(x) - f(x)| < \varepsilon \text{ for } n \geq N, \, \forall x \in S.
    \]
    Equivalently, \( \lim_{n \to \infty} \sup_{x \in S} |f_n(x) - f(x)| = 0 \).
\end{itemize}
\end{definition}

\begin{theorem}[Preservation of Continuity (24.3)]
If \( f_n \to f \) uniformly on \( S \), and each \( f_n \) is continuous at \( x_0 \in S \), then \( f \) is continuous at \( x_0 \).
\end{theorem}

\begin{proof}[Sketch of Proof]
Using an \( \varepsilon/3 \) argument, fix \( \varepsilon > 0 \). For \( f_n \to f \) uniformly, find \( n \) such that \( |f_n(x) - f(x)| < \varepsilon/3 \). By continuity of \( f_n \), there exists \( \delta > 0 \) such that \( |f_n(x_0) - f_n(x)| < \varepsilon/3 \) for \( |x - x_0| < \delta \). Combine inequalities to conclude \( |f(x_0) - f(x)| < \varepsilon \).
\end{proof}

\subsection*{Uniformly Cauchy Sequences}
\begin{definition}[Uniformly Cauchy (25.3)]
A sequence \( (f_n) \) of functions \( S \to \mathbb{R} \) is uniformly Cauchy if:
\[
\forall \varepsilon > 0, \, \exists N \in \mathbb{N} \text{ such that } |f_i(x) - f_j(x)| < \varepsilon \, \forall x \in S \text{ for } i, j \geq N.
\]
Equivalently, \( \sup_{x \in S} |f_i(x) - f_j(x)| < \varepsilon \) for \( i, j \geq N \).
\end{definition}

\begin{theorem}[Uniformly Cauchy \( \iff \) Uniform Convergence (25.4)]
A sequence \( (f_n) \) is uniformly Cauchy if and only if it converges uniformly to some \( f \).
\end{theorem}

\subsection*{Series of Functions}
\begin{definition}[Convergence of Series]
A series \( \sum_{n=1}^\infty g_n(x) \) converges (uniformly) if the sequence of partial sums \( s_k(x) = \sum_{n=1}^k g_n(x) \) converges (uniformly).
\end{definition}

\begin{theorem}[Uniform Convergence Preserves Continuity (25.5)]
If \( g_n : S \to \mathbb{R} \) are continuous and \( \sum_{n=1}^\infty g_n(x) \) converges uniformly on \( S \), then \( \sum_{n=1}^\infty g_n(x) \) is continuous.
\end{theorem}

\subsection*{Weierstrass \( M \)-Test}
\begin{theorem}[Weierstrass \( M \)-Test (25.7)]
Suppose \( M_1, M_2, \ldots \geq 0 \) and \( \sum_{k=1}^\infty M_k < \infty \). If \( |g_k(x)| \leq M_k \) for all \( x \in S \) and \( k \), then \( \sum_{k=1}^\infty g_k(x) \) converges uniformly on \( S \).
\end{theorem}

\begin{corollary}
A power series \( \sum_{k=0}^\infty a_k x^k \) converges uniformly (to a continuous function) on \( [-b, b] \) if \( b < R \), where \( R = (\limsup |a_k|^{1/k})^{-1} \).
\end{corollary}

\begin{remark}
Convergence need not be uniform on \( (-R, R) \). For example, \( \sum_{k=0}^\infty x^k = \frac{1}{1-x} \) converges on \( (-1, 1) \), but not uniformly because the partial sums are bounded while \( \frac{1}{1-x} \) is not.
\end{remark}
\section*{Limits and Differentiation (Sections 20, 28-29)}

\subsection*{Limits}
\begin{definition}[Limit (20.1, slightly modified)]
Suppose \( S \subset \mathbb{R} \), \( a \in S^- \), \( f : S \to \mathbb{R} \), and \( L \in \mathbb{R} \cup \{\pm \infty\} \). Then:
\[
\lim_{x \to a, S} f = L
\]
if \( \lim f(x_n) = L \) for any sequence \( (x_n) \subset S \) with \( \lim x_n = a \). Such sequences \( (x_n) \) exist because \( a \in S^- \).
\end{definition}

\begin{proposition}[Connection Between Limits and Continuity]
If \( a \in S \), then \( f : S \to \mathbb{R} \) is continuous at \( a \) if and only if \( \lim_{x \to a, S} f = f(a) \).
\end{proposition}

\subsection*{Common Set-Ups for Limits}
\begin{itemize}
    \item **Usual Limit**: Let \( I \) be an interval, \( a \) be interior to \( I \), and \( S = I \setminus \{a\} \). Write \( \lim_{x \to a} f \) instead of \( \lim_{x \to a, S} f \).
    \item **One-Sided Limit**: For \( S = (a, b) \), write \( \lim_{x \to a^+} f \) (right-hand limit). Define \( \lim_{x \to a^-} f \) similarly.
\end{itemize}

\subsection*{Useful Theorems About Limits}
\begin{theorem}[Equivalent Definition of Limits (20.6, Simplified)]
Suppose \( a \in S^- \). For \( f : S \to \mathbb{R} \) and \( L \in \mathbb{R} \), the following are equivalent:
\begin{enumerate}
    \item \( \lim_{x \to a, S} f = L \).
    \item For all \( \varepsilon > 0 \), there exists \( \delta > 0 \) such that \( |f(x) - L| < \varepsilon \) whenever \( x \in (a - \delta, a + \delta) \cap S \setminus \{a\} \).
\end{enumerate}
\end{theorem}

\begin{theorem}[Limit Operations (20.4)]
Suppose \( \lim_{x \to a, S} f_1 = L_1 \) and \( \lim_{x \to a, S} f_2 = L_2 \). Then:
\begin{enumerate}
    \item \( \lim_{x \to a, S} (f_1 + f_2) = L_1 + L_2 \),
    \item \( \lim_{x \to a, S} (f_1 \cdot f_2) = L_1 \cdot L_2 \),
    \item If \( L_2 \neq 0 \), \( \lim_{x \to a, S} \frac{f_1}{f_2} = \frac{L_1}{L_2} \).
\end{enumerate}
\end{theorem}

\begin{theorem}[Squeeze Theorem]
If \( f(x) \leq g(x) \leq h(x) \) for all \( x \in S \), and \( \lim_{x \to a, S} f = \lim_{x \to a, S} h = L \), then \( \lim_{x \to a, S} g = L \).
\end{theorem}

\subsection*{Differentiation}
\begin{definition}[Derivative (28.1)]
Suppose \( I \) is an open interval, \( a \in I \), and \( f : I \to \mathbb{R} \). The derivative of \( f \) at \( a \) is:
\[
f'(a) = \lim_{x \to a} \frac{f(x) - f(a)}{x - a},
\]
if the limit exists and is finite.
\end{definition}

\subsection*{Rules of Differentiation}
\begin{theorem}[Product Rule]
If \( f \) and \( g \) are differentiable at \( a \), then \( (fg)'(a) = f'(a)g(a) + f(a)g'(a) \).
\end{theorem}

\begin{theorem}[Chain Rule (28.4)]
If \( f \) is differentiable at \( a \), and \( g \) is differentiable at \( f(a) \), then \( g \circ f \) is differentiable at \( a \), with:
\[
(g \circ f)'(a) = g'(f(a)) \cdot f'(a).
\]
\end{theorem}

\subsection*{Carathéodory’s Theorem}
\begin{theorem}[Carathéodory (Exercise 28.16)]
Suppose \( I \) is an interval, \( f : I \to \mathbb{R} \). \( f \) is differentiable at \( a \in I \) if and only if there exists a function \( \phi : I \to \mathbb{R} \), continuous at \( a \), such that:
\[
f(x) - f(a) = \phi(x) \cdot (x - a), \quad \forall x \in I,
\]
and \( \phi(a) = f'(a) \).
\end{theorem}
\section*{Rules of Differentiation and Mean Value Theorem (Sections 28-29)}

\subsection*{Definition of Derivative}
\begin{definition}[Derivative (28.1)]
Suppose \( I \) is an open interval and \( a \in I \). A function \( f : I \to \mathbb{R} \) is differentiable at \( a \) if the derivative:
\[
f'(a) = \lim_{x \to a} \frac{f(x) - f(a)}{x - a}
\]
exists and is finite.
\end{definition}

\subsection*{Rules of Differentiation}
\begin{theorem}[Product Rule (28.3)]
Suppose \( f \) and \( g \) are differentiable at \( a \). Then \( fg \) is differentiable at \( a \), with:
\[
(fg)'(a) = f'(a)g(a) + f(a)g'(a).
\]
\end{theorem}

\begin{corollary}
If \( f(x) = x^m \) for \( m \in \mathbb{N} \), then:
\[
(x^m)' = mx^{m-1}.
\]
\end{corollary}

\begin{theorem}[Chain Rule (28.4)]
Suppose \( f \) is differentiable at \( a \) and \( g \) is differentiable at \( f(a) \). Then \( g \circ f \) is differentiable at \( a \), with:
\[
(g \circ f)'(a) = g'(f(a))f'(a).
\]
\end{theorem}

\subsection*{Examples of Differentiation}
\begin{enumerate}
    \item If \( f(x) = x^n \) for \( n \in \mathbb{N} \), then \( f'(x) = nx^{n-1} \).
    \item If \( f(x) = x^{-n} \) for \( n \in \mathbb{N} \), then \( f'(x) = -nx^{-n-1} \).
    \item For \( f(x) = 1/g(x) \), if \( g(a) \neq 0 \), then:
    \[
    \left( \frac{1}{g} \right)'(a) = -\frac{g'(a)}{g(a)^2}.
    \]
\end{enumerate}

\subsection*{Criterion for Extrema}
\begin{theorem}[Extrema Criterion (29.1)]
Suppose \( f \) is defined on an open interval \( I \) and has a maximum or minimum at \( x_0 \in I \). If \( f \) is differentiable at \( x_0 \), then:
\[
f'(x_0) = 0.
\]
\end{theorem}

\begin{corollary}
Suppose \( f : [a, b] \to \mathbb{R} \) is continuous and attains its maximum or minimum at \( x_0 \). Then one of the following holds:
\begin{enumerate}
    \item \( x_0 \in \{a, b\} \),
    \item \( f \) is not differentiable at \( x_0 \),
    \item \( f'(x_0) = 0 \).
\end{enumerate}
\end{corollary}

\subsection*{Rolle’s Theorem}
\begin{theorem}[Rolle’s Theorem (29.2)]
Suppose \( f : [a, b] \to \mathbb{R} \) is continuous, differentiable on \( (a, b) \), and \( f(a) = f(b) \). Then there exists \( c \in (a, b) \) such that:
\[
f'(c) = 0.
\]
\end{theorem}

\subsection*{Mean Value Theorem}
\begin{theorem}[Mean Value Theorem (29.3)]
Suppose \( f : [a, b] \to \mathbb{R} \) is continuous, differentiable on \( (a, b) \). Then there exists \( c \in (a, b) \) such that:
\[
f'(c) = \frac{f(b) - f(a)}{b - a}.
\]
\end{theorem}

\subsection*{Examples and Consequences of MVT}
\begin{corollary}[Constant Function (29.4)]
If \( f \) is differentiable on \( (a, b) \) and \( f'(x) = 0 \) for all \( x \in (a, b) \), then \( f \) is a constant function.
\end{corollary}

\begin{corollary}[Equality of Derivatives (29.5)]
If \( f \) and \( g \) are differentiable on \( (a, b) \) and \( f'(x) = g'(x) \) for all \( x \in (a, b) \), then \( f(x) - g(x) = c \) for some constant \( c \).
\end{corollary}
\section*{Rolle’s Theorem, Mean Value Theorem, and Applications (Section 29)}

\subsection*{Rolle’s Theorem}
\begin{theorem}[Rolle’s Theorem (29.2)]
Suppose \( f : [a, b] \to \mathbb{R} \) is continuous, differentiable on \( (a, b) \), and \( f(a) = f(b) \). Then there exists \( x \in (a, b) \) such that:
\[
f'(x) = 0.
\]
\end{theorem}

\begin{proof}
The function \( f \) attains its maximum and minimum on \( [a, b] \). Let \( x_0, y_0 \in [a, b] \) such that \( f(y_0) \leq f(x) \leq f(x_0) \) for all \( x \in [a, b] \). If \( f(y_0) = f(a) = f(b) = f(x_0) \), then \( f \) is constant, so \( f' = 0 \) on \( (a, b) \). Otherwise:
\begin{itemize}
    \item If \( f(x_0) > f(a) = f(b) \), then \( x_0 \in (a, b) \), and \( f'(x_0) = 0 \).
    \item If \( f(y_0) < f(a) = f(b) \), then \( y_0 \in (a, b) \), and \( f'(y_0) = 0 \).
\end{itemize}
\end{proof}

\subsection*{Mean Value Theorem}
\begin{theorem}[Mean Value Theorem (29.3)]
Suppose \( f : [a, b] \to \mathbb{R} \) is continuous, differentiable on \( (a, b) \). Then there exists \( x \in (a, b) \) such that:
\[
f'(x) = \frac{f(b) - f(a)}{b - a}.
\]
\end{theorem}

\begin{proof}
Define \( L(x) = f(a) + \frac{f(b) - f(a)}{b - a}(x - a) \) and \( g(x) = f(x) - L(x) \). The function \( g \) is continuous on \( [a, b] \), differentiable on \( (a, b) \), and \( g(a) = g(b) = 0 \). By Rolle’s Theorem, there exists \( x \in (a, b) \) such that \( g'(x) = 0 \). Thus:
\[
f'(x) = g'(x) + L'(x) = 0 + \frac{f(b) - f(a)}{b - a}.
\]
\end{proof}

\begin{example}[MVT Application]
For \( x, y \in \mathbb{R} \), \( |\sin x - \sin y| \leq |x - y| \). Apply MVT to \( f(t) = \sin t \) on \( [x, y] \): \( \exists z \in (x, y) \) such that:
\[
\frac{f(x) - f(y)}{x - y} = f'(z) = \cos z.
\]
Since \( |\cos z| \leq 1 \), \( \left| \frac{f(x) - f(y)}{x - y} \right| = |\cos z| \leq 1 \), hence \( |\sin x - \sin y| \leq |x - y| \).
\end{example}

\subsection*{Corollaries of MVT}
\begin{corollary}[Constant Functions (29.4)]
If \( f \) is differentiable on \( (a, b) \) and \( f' = 0 \) on \( (a, b) \), then \( f \) is constant.
\end{corollary}

\begin{proof}
If \( f \) is not constant, then there exist \( x < y \) such that \( f(x) \neq f(y) \). By MVT, \( \exists z \in (x, y) \) such that:
\[
f'(z) = \frac{f(y) - f(x)}{y - x} \neq 0,
\]
contradicting \( f'(z) = 0 \).
\end{proof}

\begin{corollary}[Equality of Derivatives (29.5)]
If \( f, g \) are differentiable on \( (a, b) \) and \( f' = g' \) on \( (a, b) \), then \( \exists c \in \mathbb{R} \) such that \( f(x) - g(x) = c \) for all \( x \in (a, b) \).
\end{corollary}

\begin{proof}
Define \( h(x) = f(x) - g(x) \). Then \( h' = f' - g' = 0 \). By Corollary 29.4, \( h \) is constant.
\end{proof}

\subsection*{Increasing and Decreasing Functions}
\begin{definition}[Monotonicity (29.6)]
A function \( f \) on an interval \( I \) is:
\begin{itemize}
    \item \textbf{Increasing} if \( f(x_1) \leq f(x_2) \) for \( x_1 < x_2 \),
    \item \textbf{Strictly increasing} if \( f(x_1) < f(x_2) \) for \( x_1 < x_2 \).
\end{itemize}
\end{definition}

\begin{corollary}[Monotonicity and Derivatives (29.7)]
Suppose \( f \) is differentiable on \( (a, b) \):
\begin{enumerate}
    \item \( f \) is increasing if and only if \( f' \geq 0 \) on \( (a, b) \),
    \item If \( f' > 0 \) on \( (a, b) \), then \( f \) is strictly increasing.
\end{enumerate}
\end{corollary}

\begin{example}[Bernoulli’s Inequality]
If \( n \in \mathbb{N} \) and \( x > -1 \), then:
\[
(1 + x)^n \geq 1 + nx.
\]
Let \( f(x) = (1 + x)^n - (1 + nx) \) and show \( f(x) \geq 0 \) for \( x > -1 \). By differentiating \( f(x) \), we conclude \( f(x) \) is increasing and achieves its minimum at \( x = 0 \), where \( f(0) = 0 \).
\end{example}
\section*{Differentiating Inverse Functions and Integration (Sections 29, 32)}

\subsection*{Differentiating Inverse Functions}
\begin{theorem}[Derivative of an Inverse Function (29.9)]
Suppose \( I \) is an interval, \( f : I \to \mathbb{R} \) is a continuous, strictly monotone function. Let \( J = f(I) \), and \( g = f^{-1} : J \to I \). If \( f \) is differentiable at \( c \in I \), and \( f'(c) \neq 0 \), then \( g \) is differentiable at \( d = f(c) \), and:
\[
g'(d) = \frac{1}{f'(c)} = \frac{1}{f'(g(d))}.
\]
\end{theorem}

\begin{proof}[Proof Sketch]
Using Carathéodory’s Theorem:
\[
f(x) - f(c) = \phi(x)(x - c), \quad \phi(c) = f'(c),
\]
where \( \phi \) is continuous at \( c \). For \( y = f(g(y)) \), differentiate both sides to find \( g'(d) = 1 / \phi(g(d)) \).
\end{proof}

\subsection*{Derivatives of Rational Powers}
\begin{example}[Derivative of Rational Powers]
Let \( f(x) = x^n \) (strictly increasing on \( (0, \infty) \)) with \( f'(x) = nx^{n-1} \). The inverse function is \( g(y) = y^{1/n} \). For \( y > 0 \):
\[
g'(y) = \frac{1}{f'(g(y))} = \frac{1}{n(y^{1/n})^{n-1}} = \frac{1}{n} y^{1/n - 1}.
\]
If \( n \in \mathbb{Z} \) is odd, extend \( f, g \) to \( \mathbb{R} \). Then \( g'(y) = \frac{1}{n} y^{1/n - 1} \) for \( y < 0 \).
\end{example}

\begin{example}[Derivative of \( h(x) = x^r \), \( r \in \mathbb{Q} \)]
Write \( r = m/n \), \( h(x) = x^{m/n} \). Use the chain rule:
\[
h'(x) = \frac{m}{n} x^{r-1}.
\]
\end{example}

\subsection*{Inverse Trigonometric Functions}
\begin{example}[Arcsine]
For \( f(x) = \sin x \) on \( [-\pi/2, \pi/2] \), \( g = \arcsin : [-1, 1] \to [-\pi/2, \pi/2] \). Since \( f'(x) = \cos x \), for \( y \in (-1, 1) \):
\[
(\arcsin y)' = \frac{1}{\sqrt{1-y^2}}.
\]
\end{example}

\begin{example}[Arctangent]
For \( f(x) = \tan x \) on \( (-\pi/2, \pi/2) \), \( g = \arctan : \mathbb{R} \to (-\pi/2, \pi/2) \). Since \( f'(x) = 1 + x^2 \), for \( y \in \mathbb{R} \):
\[
(\arctan y)' = \frac{1}{1 + y^2}.
\]
\end{example}

\subsection*{Integration: Concepts and Definitions}
\begin{definition}[Darboux Sums]
Let \( f : [a, b] \to \mathbb{R} \) be bounded.
\begin{itemize}
    \item Partition \( P = \{t_0, t_1, \ldots, t_n\} \) of \( [a, b] \) gives subintervals \( [t_{k-1}, t_k] \).
    \item Lower Darboux sum:
    \[
    L(f, P) = \sum_{k=1}^n m(f, [t_{k-1}, t_k])(t_k - t_{k-1}),
    \]
    where \( m(f, [t_{k-1}, t_k]) = \inf_{x \in [t_{k-1}, t_k]} f(x) \).
    \item Upper Darboux sum:
    \[
    U(f, P) = \sum_{k=1}^n M(f, [t_{k-1}, t_k])(t_k - t_{k-1}),
    \]
    where \( M(f, [t_{k-1}, t_k]) = \sup_{x \in [t_{k-1}, t_k]} f(x) \).
\end{itemize}
\end{definition}

\begin{definition}[Integrability]
\( f \) is integrable on \( [a, b] \) if:
\[
\sup_P L(f, P) = \inf_P U(f, P),
\]
denoted \( \int_a^b f(x) dx \).
\end{definition}

\subsection*{Examples of Integrability}
\begin{example}[Constant Function]
If \( f(x) = c \), then:
\[
\int_a^b f(x) dx = c(b-a).
\]
\end{example}

\begin{example}[Discontinuous Function]
Let \( g(x) = 1 \) if \( x \in \mathbb{Q} \), \( g(x) = 0 \) otherwise. Then:
\[
\sup_P L(g, P) = 0, \quad \inf_P U(g, P) = 1.
\]
Since \( \sup_P L \neq \inf_P U \), \( g \) is not integrable.
\end{example}

\begin{example}[Linear Function]
If \( h(x) = x \), then:
\[
\int_0^b h(x) dx = \frac{b^2}{2}.
\]
\end{example}
\section*{Darboux Sums, Integrability, and Riemann Integration (Sections 32-33)}

\subsection*{Darboux Sums and Integrals}
\begin{definition}[Darboux Sums]
Let \( f : [a, b] \to \mathbb{R} \) be bounded. For a partition \( P = \{a = t_0 < t_1 < \cdots < t_n = b\} \):
\begin{itemize}
    \item Lower Darboux sum:
    \[
    L(f, P) = \sum_{k=1}^n m(f, [t_{k-1}, t_k])(t_k - t_{k-1}),
    \]
    where \( m(f, [t_{k-1}, t_k]) = \inf_{x \in [t_{k-1}, t_k]} f(x) \).
    \item Upper Darboux sum:
    \[
    U(f, P) = \sum_{k=1}^n M(f, [t_{k-1}, t_k])(t_k - t_{k-1}),
    \]
    where \( M(f, [t_{k-1}, t_k]) = \sup_{x \in [t_{k-1}, t_k]} f(x) \).
\end{itemize}
\end{definition}

\begin{definition}[Integrability]
The lower Darboux integral is \( L(f) = \sup_P L(f, P) \), and the upper Darboux integral is \( U(f) = \inf_P U(f, P) \). \( f \) is integrable if \( L(f) = U(f) \), denoted:
\[
\int_a^b f = L(f) = U(f).
\]
\end{definition}

\begin{theorem}[32.4]
If \( f : [a, b] \to \mathbb{R} \) is bounded, then \( L(f) \leq U(f) \).
\end{theorem}

\subsection*{Integrals: Example}
\begin{example}[Linear Function]
Is \( h(x) = x \) integrable on \( [0, b] \)? Compute \( \int_0^b h(x) dx \).

For \( P = \{0, \frac{b}{n}, \frac{2b}{n}, \ldots, b\} \):
\[
L(h, P_n) = \frac{b^2}{2}\left(1 - \frac{1}{n}\right), \quad U(h, P_n) = \frac{b^2}{2}.
\]
Thus:
\[
L(h) \geq \sup_n L(h, P_n) = \frac{b^2}{2}, \quad U(h) \leq \lim_n U(h, P_n) = \frac{b^2}{2}.
\]
Since \( L(h) = U(h) \), \( h(x) \) is integrable with:
\[
\int_0^b h(x) dx = \frac{b^2}{2}.
\]
\end{example}

\subsection*{Criterion for Integrability}
\begin{theorem}[32.5]
A bounded function \( f : [a, b] \to \mathbb{R} \) is integrable if and only if for all \( \varepsilon > 0 \), there exists a partition \( P \) such that:
\[
U(f, P) - L(f, P) < \varepsilon.
\]
\end{theorem}

\subsection*{Monotone and Continuous Functions}
\begin{theorem}[33.1]
Any monotone function on \( [a, b] \) is integrable.
\end{theorem}

\begin{theorem}[33.2]
Any continuous function on \( [a, b] \) is integrable.
\end{theorem}

\subsection*{Mesh of a Partition}
\begin{definition}[Mesh (32.6)]
The mesh of a partition \( P = \{t_0, t_1, \ldots, t_n\} \) is:
\[
\text{mesh}(P) = \max_{1 \leq k \leq n} (t_k - t_{k-1}).
\]
\end{definition}

\begin{theorem}[32.7]
A bounded \( f : [a, b] \to \mathbb{R} \) is integrable if and only if for all \( \varepsilon > 0 \), there exists \( \delta > 0 \) such that:
\[
U(f, P) - L(f, P) < \varepsilon \quad \text{whenever mesh}(P) < \delta.
\]
\end{theorem}

\subsection*{Riemann Integration}
\begin{definition}[Riemann Integral (32.8)]
Let \( f : [a, b] \to \mathbb{R} \) be bounded. For a partition \( P = \{t_0, t_1, \ldots, t_n\} \) and \( x_k \in [t_{k-1}, t_k] \), define the Riemann sum:
\[
S = \sum_{k=1}^n f(x_k)(t_k - t_{k-1}).
\]
\( f \) is Riemann integrable if there exists \( r \in \mathbb{R} \) such that for all \( \varepsilon > 0 \), there exists \( \delta > 0 \) such that:
\[
|S - r| < \varepsilon \quad \text{whenever mesh}(P) < \delta.
\]
\end{definition}

\begin{theorem}[32.9]
A bounded function \( f : [a, b] \to \mathbb{R} \) is Riemann integrable if and only if it is Darboux integrable. In this case:
\[
\int_a^b f = R\int_a^b f.
\]
\end{theorem}
\section*{Integrability and Riemann Integration (Sections 32-33)}

\subsection*{Monotone and Continuous Functions}
\begin{theorem}[Integrability Criterion (32.5)]
A bounded function \( f : [a, b] \to \mathbb{R} \) is integrable if and only if:
\[
\forall \varepsilon > 0, \exists \text{ a partition } P \text{ such that } U(f, P) - L(f, P) < \varepsilon.
\]
\end{theorem}

\begin{theorem}[Monotone Functions Are Integrable (33.1)]
Any monotone function on \( [a, b] \) is integrable.
\end{theorem}

\begin{proof}
Assume \( f \) is increasing. Fix \( \varepsilon > 0 \). Choose \( n \in \mathbb{N} \) such that:
\[
\frac{(f(b) - f(a))(b - a)}{n} < \varepsilon.
\]
Consider the partition \( P \) with \( t_k = a + kh \) for \( 0 \leq k \leq n \), where \( h = \frac{b-a}{n} \). Then:
\[
U(f, P) - L(f, P) = h \sum_{k=1}^n (f(t_k) - f(t_{k-1})) = \frac{(f(b) - f(a))(b - a)}{n} < \varepsilon.
\]
\end{proof}

\begin{theorem}[Continuous Functions Are Integrable (33.2)]
Any continuous function on \( [a, b] \) is integrable.
\end{theorem}

\begin{proof}
Fix \( \varepsilon > 0 \). By uniform continuity, \( \exists \delta > 0 \) such that \( |f(x) - f(y)| < \frac{\varepsilon}{b-a} \) whenever \( |x - y| < \delta \). Choose \( n \in \mathbb{N} \) such that \( h = \frac{b-a}{n} < \delta \), and partition \( P \) with \( t_k = a + kh \). Then:
\[
M(f, [t_{k-1}, t_k]) - m(f, [t_{k-1}, t_k]) < \frac{\varepsilon}{b-a}.
\]
Thus:
\[
U(f, P) - L(f, P) < hn \cdot \frac{\varepsilon}{b-a} = \varepsilon.
\]
\end{proof}

\subsection*{Mesh of a Partition}
\begin{definition}[Mesh (32.6)]
The mesh of a partition \( P = \{t_0, t_1, \ldots, t_n\} \) is:
\[
\text{mesh}(P) = \max_{1 \leq k \leq n} (t_k - t_{k-1}).
\]
\end{definition}

\begin{theorem}[Integrability and Mesh (32.7)]
A bounded function \( f : [a, b] \to \mathbb{R} \) is integrable if and only if:
\[
\forall \varepsilon > 0, \exists \delta > 0 \text{ such that } U(f, P) - L(f, P) < \varepsilon \text{ whenever mesh}(P) < \delta.
\]
\end{theorem}

\subsection*{Riemann Integration}
\begin{definition}[Riemann Integral (32.8)]
Let \( f : [a, b] \to \mathbb{R} \) be bounded. For a partition \( P = \{t_0, t_1, \ldots, t_n\} \) and \( x_k \in [t_{k-1}, t_k] \), define the Riemann sum:
\[
S = \sum_{k=1}^n f(x_k)(t_k - t_{k-1}).
\]
\( f \) is Riemann integrable if:
\[
\exists r \in \mathbb{R} \text{ such that } \forall \varepsilon > 0, \exists \delta > 0 \text{ such that } |S - r| < \varepsilon \text{ whenever mesh}(P) < \delta.
\]
\end{definition}

\begin{theorem}[Equivalence of Riemann and Darboux Integrability (32.9)]
A bounded \( f : [a, b] \to \mathbb{R} \) is Riemann integrable if and only if it is Darboux integrable. In this case:
\[
R\int_a^b f = \int_a^b f.
\]
\end{theorem}

\subsection*{Properties of Integrable Functions}
\begin{proposition}[Exercise 32.7]
If \( f \) is integrable on \( [a, b] \), and \( f = g \) except at finitely many points, then \( g \) is integrable on \( [a, b] \) and:
\[
\int_a^b f = \int_a^b g.
\]
\end{proposition}

\begin{remark}
The statement fails if the set of exceptions is countably infinite. For example:
\[
f(x) = 0, \quad g(x) =
\begin{cases}
1 & x \in \mathbb{Q}, \\
0 & x \notin \mathbb{Q}.
\end{cases}
\]
\( f \) is integrable, but \( g \) is not.
\end{remark}
\section*{Properties of Integrals and Convergence Theorems (Section 33)}

\subsection*{Properties of Integrals}
\begin{theorem}[Linearity and Comparison of Integrals (33.3, 33.4(i))]
Suppose \( f, g \) are integrable on \( [a, b] \), and \( c \in \mathbb{R} \). Then:
\begin{enumerate}
    \item \( cf \) is integrable, and:
    \[
    \int_a^b cf = c \int_a^b f.
    \]
    \item \( f + g \) is integrable, and:
    \[
    \int_a^b (f + g) = \int_a^b f + \int_a^b g.
    \]
    \item If \( f \geq g \), then:
    \[
    \int_a^b f \geq \int_a^b g.
    \]
\end{enumerate}
\end{theorem}

\begin{theorem}[Triangle Inequality for Integrals (33.5)]
If \( f \) is integrable on \( [a, b] \), then \( |f| \) is integrable, and:
\[
\left| \int_a^b f \right| \leq \int_a^b |f|.
\]
\end{theorem}

\subsection*{Integrability of Products and Piecewise Functions}
\begin{proposition}
If \( f \) is integrable on \( [a, b] \), then \( f^2 \) is integrable.
\end{proposition}

\begin{corollary}
If \( f, g \) are integrable on \( [a, b] \), then \( fg \) is integrable.
\end{corollary}

\begin{proof}
Express \( fg \) as:
\[
fg = \frac{1}{4} \left( (f + g)^2 - (f - g)^2 \right).
\]
Since \( f + g \) and \( f - g \) are integrable, their squares are integrable, and hence \( fg \) is integrable.
\end{proof}

\begin{theorem}[Piecewise Monotone and Continuous Functions (33.8)]
Suppose \( f : [a, b] \to \mathbb{R} \) is either:
\begin{enumerate}
    \item Piecewise monotone and bounded, or
    \item Piecewise continuous.
\end{enumerate}
Then \( f \) is integrable.
\end{theorem}

\begin{proof}
Partition \( [a, b] \) such that \( f \) is monotone or uniformly continuous on each subinterval. On each subinterval, \( f \) is integrable. By additivity of the integral, \( f \) is integrable on \( [a, b] \).
\end{proof}

\subsection*{Convergence and Interchange of Limits and Integrals}
\begin{proposition}
Suppose \( (f_n) \) is a sequence of integrable functions on \( [a, b] \) that converges uniformly to \( f \). Then \( f \) is integrable, and:
\[
\int_a^b f = \lim_{n \to \infty} \int_a^b f_n.
\]
\end{proposition}

\subsection*{Convergence Theorems}
\begin{theorem}[Bounded Convergence (33.11)]
Suppose \( (f_n) \) are integrable on \( [a, b] \), \( |f_n| \leq M \) for all \( n \), \( f_n \to f \) pointwise on \( [a, b] \), and \( f \) is integrable. Then:
\[
\lim_{n \to \infty} \int_a^b f_n = \int_a^b f.
\]
\end{theorem}

\begin{theorem}[Monotone Convergence (33.12)]
Suppose \( (f_n) \) are integrable on \( [a, b] \), \( f_1 \leq f_2 \leq \cdots \), \( f_n \to f \) pointwise on \( [a, b] \), and \( f \) is integrable. Then:
\[
\lim_{n \to \infty} \int_a^b f_n = \int_a^b f.
\]
\end{theorem}

\begin{example}[Application of Monotone Convergence]
Let \( f_n(x) = \frac{1}{1 + nx^3} \) on \( [0, 1] \). Then:
\[
\int_0^1 f_n(x) dx \to \int_0^1 f(x) dx = 0,
\]
where \( f(x) = \begin{cases} 1, & x = 0, \\ 0, & x \in (0, 1]. \end{cases} \)
\end{example}
\section*{Fundamental Theorems of Calculus and Change of Variable (Section 34)}

\subsection*{Fundamental Theorem of Calculus I}
\begin{theorem}[Fundamental Theorem of Calculus I (34.1)]
Suppose \( g : [a, b] \to \mathbb{R} \) is continuous, differentiable on \( (a, b) \), and \( g' \) is integrable on \( [a, b] \). Then:
\[
\int_a^b g'(x) \, dx = g(b) - g(a).
\]
\end{theorem}

\begin{example}
Compute \( \int_a^b x^n \, dx \). Use \( g(x) = \frac{x^{n+1}}{n+1} \), so:
\[
\int_a^b x^n \, dx = \frac{b^{n+1} - a^{n+1}}{n+1}.
\]
\end{example}

\begin{proof}
Partition \( P = \{a = t_0 < t_1 < \ldots < t_n = b\} \). By the Mean Value Theorem:
\[
g'(x_k) = \frac{g(t_k) - g(t_{k-1})}{t_k - t_{k-1}},
\]
for some \( x_k \in (t_{k-1}, t_k) \). Then:
\[
L(g', P) \leq \sum_{k=1}^n g'(x_k)(t_k - t_{k-1}) = g(b) - g(a) \leq U(g', P).
\]
Thus \( \int_a^b g'(x) \, dx = g(b) - g(a) \).
\end{proof}

\subsection*{Integration by Parts}
\begin{theorem}[Integration by Parts (34.2)]
Suppose \( u, v : [a, b] \to \mathbb{R} \) are continuous, differentiable on \( (a, b) \), and \( u', v' \) are integrable on \( [a, b] \). Then:
\[
\int_a^b u(x)v'(x) \, dx + \int_a^b u'(x)v(x) \, dx = u(b)v(b) - u(a)v(a).
\]
\end{theorem}

\begin{example}
Compute \( \int_0^\pi x \cos x \, dx \). Let \( u(x) = x \), \( v'(x) = \cos x \):
\[
\int_0^\pi x \cos x \, dx = x \sin x \big|_0^\pi - \int_0^\pi \sin x \, dx = 0 - (-2) = -2.
\]
\end{example}

\subsection*{Fundamental Theorem of Calculus II}
\begin{theorem}[Fundamental Theorem of Calculus II (34.3)]
Suppose \( f : [a, b] \to \mathbb{R} \) is integrable. Define:
\[
F(x) = \int_a^x f(t) \, dt.
\]
If \( f \) is continuous at \( c \), then \( F \) is differentiable at \( c \), with \( F'(c) = f(c) \).
\end{theorem}

\begin{example}
Let \( G(x) = \int_{x^2}^2 \sin(t^2) \, dt \). Then \( G'(x) = -2x \sin(x^4) \) by the Chain Rule.
\end{example}

\begin{proof}
Let \( F(x) = \int_a^x f(t) \, dt \). Then:
\[
F'(c) = \lim_{x \to c} \frac{F(x) - F(c)}{x - c} = \lim_{x \to c} \frac{\int_c^x f(t) \, dt}{x - c}.
\]
Since \( f \) is continuous at \( c \), \( |f(t) - f(c)| \leq \varepsilon \) for \( |t - c| < \delta \), and thus:
\[
\lim_{x \to c} \frac{\int_c^x f(t) \, dt}{x - c} = f(c).
\]
\end{proof}

\subsection*{Change of Variable in Integrals}
\begin{theorem}[Change of Variable (34.4)]
Suppose \( u : J \to I \), \( u' \) is continuous, and \( f : I \to \mathbb{R} \) is continuous. Then for \( a, b \in J \):
\[
\int_a^b f(u(x)) u'(x) \, dx = \int_{u(a)}^{u(b)} f(t) \, dt.
\]
\end{theorem}

\begin{example}
Compute \( \int_1^4 \frac{\sin(\sqrt{x})}{\sqrt{x}} \, dx \). Let \( u(x) = \sqrt{x} \), then \( u'(x) = \frac{1}{2\sqrt{x}} \):
\[
\int_1^4 \frac{\sin(\sqrt{x})}{\sqrt{x}} \, dx = \int_1^2 2 \sin t \, dt = 2(-\cos t) \big|_1^2 = 2(\cos 1 - \cos 2).
\]
\end{example}
\section*{Interchanging Integration, Differentiation, and Power Series (Section 26)}

\subsection*{Interchanging Integration with Limits and Sums}
\begin{proposition}[Exercise 33.9, Lecture 32]
Suppose \( (f_n) \) is a sequence of integrable functions on \( [a, b] \) converging uniformly to \( f \). Then:
\[
\lim_{n \to \infty} \int_a^b f_n = \int_a^b f.
\]
\end{proposition}

\begin{corollary}
If \( g_n \) are integrable on \( [a, b] \), and \( f = \sum_{n=0}^\infty g_n \) converges uniformly, then \( f \) is integrable, and:
\[
\int_a^b f = \sum_{n=0}^\infty \int_a^b g_n.
\]
\end{corollary}

\subsection*{Interchanging Differentiation with Limits and Sums}
\begin{example}
Let \( f_n(x) = \frac{1}{n} \sin(n^2 x) \). Then \( f_n \to 0 \) uniformly on \( \mathbb{R} \). However:
\[
f_n'(x) = n \cos(n^2 x).
\]
If \( x = \frac{p}{q} \pi \), then \( f_n'(x) \) does not converge, even pointwise.
\end{example}

\subsection*{Power Series and Radius of Convergence}
\begin{definition}[Radius of Convergence]
For a power series \( \sum_{n=0}^\infty a_n x^n \), let:
\[
\beta = \limsup_{n \to \infty} |a_n|^{1/n}.
\]
The radius of convergence is \( R = \frac{1}{\beta} \).
\end{definition}

\begin{theorem}[Uniform Convergence (26.1)]
The series \( \sum_{n=0}^\infty a_n x^n \) converges uniformly on \( [-R_1, R_1] \) for \( R_1 < R \).
\end{theorem}

\begin{corollary}
The series \( \sum_{n=0}^\infty a_n x^n \) converges to a continuous function on \( (-R, R) \).
\end{corollary}

\subsection*{Differentiation and Integration of Power Series}
\begin{lemma}[Differentiation and Integration (26.3)]
If \( \sum_{n=0}^\infty a_n x^n \) has radius of convergence \( R \), then:
\begin{enumerate}
    \item \( \sum_{n=1}^\infty n a_n x^{n-1} \) has radius of convergence \( R \),
    \item \( \sum_{n=0}^\infty \frac{a_n}{n+1} x^{n+1} \) has radius of convergence \( R \).
\end{enumerate}
\end{lemma}

\begin{theorem}[Integration of Power Series (26.4)]
Suppose \( f(t) = \sum_{n=0}^\infty a_n t^n \) has radius of convergence \( R \). Then for \( |x| < R \):
\[
\int_0^x f(t) \, dt = \sum_{n=0}^\infty \frac{a_n}{n+1} x^{n+1}.
\]
\end{theorem}

\begin{example}
For \( f(t) = \frac{1}{1-t} = \sum_{n=0}^\infty t^n \) (\( R = 1 \)):
\[
-\ln(1-x) = \int_0^x f(t) \, dt = \sum_{n=0}^\infty \frac{x^{n+1}}{n+1}, \quad |x| < 1.
\]
\end{example}

\begin{theorem}[Differentiation of Power Series (26.5)]
Suppose \( f(t) = \sum_{n=0}^\infty a_n t^n \) has radius of convergence \( R \). Then for \( |t| < R \):
\[
f'(t) = \sum_{n=1}^\infty n a_n t^{n-1}.
\]
\end{theorem}

\subsection*{Abel’s Theorem}
\begin{theorem}[Abel’s Theorem (26.6)]
Suppose \( f(x) = \sum_{n=0}^\infty a_n x^n \) has radius of convergence \( R > 0 \). If the series converges at \( R \) (or \( -R \)), then \( f \) is continuous at \( R \) (or \( -R \)).
\end{theorem}
\section*{Abel’s Theorem, Convexity, and Inequalities (Section 26)}

\subsection*{Abel Summation Theorem}
\begin{theorem}[Abel’s Theorem (26.6)]
Let \( f(x) = \sum_{n=0}^\infty a_n x^n \) have radius of convergence \( R > 0 \). If the series converges at \( R \) (or \( -R \)), then \( f \) is continuous at \( R \) (or \( -R \).
\end{theorem}

\begin{example}
\[
1 - \frac{1}{2} + \frac{1}{3} - \cdots = \ln 2.
\]
Let \( g(t) = \frac{1}{1+t} = \sum_{n=0}^\infty (-1)^n t^n \) (\( R = 1 \)), and \( f(x) = \sum_{k=1}^\infty \frac{(-1)^{k-1}}{k} x^k \). The series diverges at \( -1 \) but converges at \( 1 \):
\[
\ln(1+x) = \int_0^x g(t) \, dt = \sum_{k=1}^\infty \frac{(-1)^{k-1}}{k} x^k.
\]
Thus:
\[
f(1) = \sum_{k=1}^\infty \frac{(-1)^{k-1}}{k} = \ln 2.
\]
\end{example}

\subsection*{Alternating Series}
\begin{proposition}[Alternating Series Test]
Suppose \( a_1 \geq a_2 \geq \cdots \geq 0 \). Then:
\[
\sum_{k=1}^\infty (-1)^{k-1} a_k = a_1 - a_2 + a_3 - \cdots
\]
converges if and only if \( \lim_{k \to \infty} a_k = 0 \).
\end{proposition}

\begin{example}
\[
1 - \frac{1}{3} + \frac{1}{5} - \cdots = \frac{\pi}{4}.
\]
Let \( f(x) = \arctan x \), so \( f'(x) = \frac{1}{1+x^2} \). For \( |x| < 1 \):
\[
f'(x) = \sum_{n=0}^\infty (-1)^n x^{2n}, \quad f(x) = \int_0^x f'(t) \, dt = \sum_{n=0}^\infty \frac{(-1)^n}{2n+1} x^{2n+1}.
\]
At \( x = 1 \):
\[
\frac{\pi}{4} = \arctan 1 = \sum_{n=0}^\infty \frac{(-1)^n}{2n+1}.
\]
\end{example}

\subsection*{Convex Functions}
\begin{definition}[Convexity]
A function \( f \) on \( I \) is convex if:
\[
f\left(\frac{x+y}{2}\right) \leq \frac{f(x) + f(y)}{2}, \quad \forall x, y \in I.
\]
\end{definition}

\begin{proposition}
If \( f \) is convex, then:
\[
f((1-t)x + ty) \leq (1-t)f(x) + tf(y), \quad \forall t \in (0, 1).
\]
\end{proposition}

\subsection*{Criteria for Convexity}
\begin{proposition}
If \( f \) is differentiable on \( I \), and \( f' \) is increasing, then \( f \) is convex.
\end{proposition}

\begin{corollary}
If \( f \) is twice differentiable on \( I \), and \( f'' \geq 0 \), then \( f \) is convex.
\end{corollary}

\begin{example}
\begin{itemize}
    \item \( f(x) = e^x \) is convex on \( \mathbb{R} \) because \( f''(x) = e^x > 0 \).
    \item \( g(x) = \ln x \) is concave on \( (0, \infty) \) because \( g''(x) = -\frac{1}{x^2} < 0 \).
\end{itemize}
\end{example}

\subsection*{Inequalities}
\begin{theorem}[Jensen’s Inequality]
If \( f \) is convex on \( I \), \( x_1, \ldots, x_n \in I \), \( t_1, \ldots, t_n \geq 0 \), and \( \sum_{i=1}^n t_i = 1 \), then:
\[
f\left(\sum_{i=1}^n t_i x_i\right) \leq \sum_{i=1}^n t_i f(x_i).
\]
\end{theorem}

\begin{proposition}[Arithmetic-Geometric Means Inequality]
If \( x_1, \ldots, x_n > 0 \), \( t_1, \ldots, t_n > 0 \), and \( \sum_{i=1}^n t_i = 1 \), then:
\[
\sum_{i=1}^n t_i x_i \geq \prod_{i=1}^n x_i^{t_i}.
\]
\end{proposition}

\begin{corollary}[Special Case]
If \( x_1, \ldots, x_n > 0 \), then:
\[
\frac{x_1 + \cdots + x_n}{n} \geq \sqrt[n]{x_1 \cdots x_n}.
\]
\end{corollary}
\section*{Convexity, Inequalities, and Nowhere Differentiable Functions (Section 36)}

\subsection*{Jensen’s Inequality for Convex Functions}
\begin{theorem}[Jensen’s Inequality]
Let \( f \) be a convex function on an interval \( I \), and let \( x_1, \ldots, x_n \in I \) with \( t_1, \ldots, t_n \geq 0 \) and \( \sum_{i=1}^n t_i = 1 \). Then:
\[
f\left(\sum_{i=1}^n t_i x_i\right) \leq \sum_{i=1}^n t_i f(x_i).
\]
If \( f \) is concave, the inequality is reversed.
\end{theorem}

\subsection*{Inequalities Between Means}
\begin{proposition}[Power Mean Inequality]
Suppose \( r > 1 \) and \( x_1, \ldots, x_n \geq 0 \). Then:
\[
\frac{x_1 + \cdots + x_n}{n} \leq \left(\frac{x_1^r + \cdots + x_n^r}{n}\right)^{1/r}.
\]
If \( r = 2 \), this gives the inequality between arithmetic and quadratic means:
\[
\frac{x_1 + \cdots + x_n}{n} \leq \sqrt{\frac{x_1^2 + \cdots + x_n^2}{n}}.
\]
\end{proposition}

\begin{proof}
On \( [0, \infty) \), \( f(x) = x^r \) is convex because \( f'(x) = rx^{r-1} \) is increasing. Apply Jensen’s Inequality with \( t_i = \frac{1}{n} \):
\[
\left(\frac{x_1 + \cdots + x_n}{n}\right)^r \leq \frac{x_1^r + \cdots + x_n^r}{n}.
\]
Taking the \( r \)-th root gives the result.
\end{proof}

\subsection*{Arithmetic and Harmonic Means}
\begin{proposition}[Arithmetic-Harmonic Mean Inequality]
If \( x_1, \ldots, x_n > 0 \), then:
\[
\frac{x_1 + \cdots + x_n}{n} \geq \frac{n}{\frac{1}{x_1} + \cdots + \frac{1}{x_n}}.
\]
\end{proposition}

\begin{proof}
Let \( g(x) = \frac{1}{x} \), which is convex on \( (0, \infty) \). Let \( y_i = \frac{1}{x_i} \). By Jensen’s Inequality with \( t_i = \frac{1}{n} \):
\[
\frac{1}{n} \sum_{i=1}^n g(y_i) = \frac{\frac{1}{x_1} + \cdots + \frac{1}{x_n}}{n} \geq g\left(\frac{1}{n} \sum_{i=1}^n y_i\right) = \frac{n}{x_1 + \cdots + x_n}.
\]
\end{proof}

\subsection*{Nowhere Differentiable Functions}
\begin{proposition}
There exists a bounded, uniformly continuous function \( f : \mathbb{R} \to \mathbb{R} \) that is differentiable nowhere.
\end{proposition}

\begin{proof}[Sketch]
Construct a \( 1 \)-periodic function \( f(x) = \sum_{k=0}^\infty 8^{-k}s(64^k x) \), where \( s(x) \) is the sawtooth function:
\[
s(x) = \phi(x - \lfloor x \rfloor), \quad \phi(t) = \min\{t, 1-t\}.
\]
\begin{enumerate}
    \item \( f \) is bounded and uniformly continuous by the Weierstrass \( M \)-test.
    \item For any \( x, \delta > 0, A > 0 \), there exists \( y \) with \( |x - y| \leq \delta \) and \( |f(x) - f(y)| \geq A|x - y| \), showing \( f \) is nowhere differentiable.
\end{enumerate}
\end{proof}

\subsection*{Infinite Primes and Divergence of Series}
\begin{theorem}
Let \( p_1 < p_2 < \cdots \) be the increasing sequence of prime numbers. Then:
\[
\sum_{n=1}^\infty \frac{1}{p_n} \quad \text{diverges}.
\]
\end{theorem}

\begin{proof}
Assume \( \sum_{n=1}^\infty \frac{1}{p_n} \) converges. Let \( \alpha = \sum_{n=1}^\infty \frac{1}{p_{2n}^2} < 1 \), and \( \beta = \sum_{n=K+1}^\infty \frac{1}{p_n} < 1 - \alpha \) for some \( K \). Consider \( N \) such that \( N > 2K/(1 - \alpha - \beta) \). Counting arguments on \( \{1, 2, \ldots, N\} \) lead to a contradiction.
\end{proof}
\end{document}